.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                5-%
.pc 1
.ce
.sh
Section 5
.qs
.sp
.ce
.sh
CP/M 2 System Interface
.qs
.tc 5  CP/M 2 System Interface
.sp 2
.he CP/M Operating System Manual                    5.1  Introduction
.tc    5.1  Introduction
5.1  Introduction
.pp 5
This chapter describes CP/M (release 2) system organization including the
structure of memory and system entry points.  This section provides
the information you need to write programs that operate under CP/M and
that use the peripheral and disk I/O facilities of the system.
.pp
CP/M is logically divided into four parts, called the Basic Input/Output 
System (BIOS), the Basic Disk Operating System (BDOS), the Console Command
Processor (CCP), and the Transient Program Area (TPA).  The BIOS is a
hardware-dependent module that defines the exact low level interface with a
particular computer system that is necessary for peripheral device I/O.
Although a standard BIOS is supplied by Digital Research, explicit
instructions are provided for field reconfiguration of the BIOS to match
nearly any hardware environment, see Section 6.
.pp
The BIOS and BDOS are
logically combined into a single module with a common entry point and
referred to as the FDOS.  The CCP is a distinct program that uses the FDOS to
provide a human-oriented interface with the information that is cataloged on
the back-up storage device.  The TPA is an area of memory, 
not used by the FDOS and CCP, where various nonresident operating
system commands and user programs are executed.  The lower portion of memory
is reserved for system information and is detailed in later sections.  Memory
organization of the CP/M system is shown in Figure 5-1.
.sp 3
.nf
                  High
                  Memory     FDOS (BDOS+BIOS)
                  FBASE:

                                   CCP
                  CBASE:

                                   TPA
                  TBASE:

                             System Parameters
                  BOOT:
.sp 2
.ce
.sh
Figure 5-1.  CP/M Memory Organization
.qs
.fi
.sp 2
.pp
The exact memory addresses corresponding to BOOT, TBASE, CBASE, and FBASE
vary from version to version and are described fully in Section 6.  All
standard CP/M versions assume BOOT=0000H, which is the base of
random access memory.  The machine code found at location BOOT performs a
system warm start, which loads and initializes the programs and variables
necessary to return control to the CCP.  Thus, transient programs need only
jump to location BOOT to return control to CP/M at the command level.
Further, the standard versions assume TBASE=BOOT+0100H, which is normally
location 0100H.  The principal entry point to the FDOS is at location
BOOT+0005H (normally 0005H) where a jump to FBASE is found.  The address
field at BOOT+0006H (normally 0006H) contains the value of FBASE and can be
used to determine the size of available memory, assuming that the CCP is
being overlaid by a transient program.
.pp
Transient programs are loaded into the TPA and executed as follows.  The
operator communicates with the CCP by typing command lines following each
prompt.  Each command line takes one of the following forms:
.sp
.nf
.in 8
command
command file1
command file1 file2
.fi
.in 0
.sp
where command is either a built-in function, such as DIR or TYPE, or the name
of a transient command or program.  If the command is a built-in function of
CP/M, it is executed immediately.  Otherwise, the CCP searches the currently
addressed disk for a file by the name
.sp
.ti 8
command.COM
.pp
If the file is found, it is assumed to be a memory image of a program that
executes in the TPA and thus implicity originates at TBASE in memory.  The
CCP loads the COM file from the disk into memory starting at TBASE and can
extend up to CBASE.
.pp
If the command is followed by one or two file specifications, the CCP prepares
one or two File Control Block (FCB) names in the system 
parameter area.  These optional FCBs are in the form necessary to 
access files through the FDOS and are described in Section 5.2.
.pp
The transient program receives control from the CCP and begins 
execution, using the I/O facilities of the FDOS.  The transient 
program is called from the CCP.  Thus, it can simply return to the CCP upon
completion of its processing, or can jump to BOOT to pass control back to
CP/M.  In the first case, the transient program must not use memory above
CBASE, while in the latter case, memory up through FBASE-1 can be used.
.pp
The transient program can use the CP/M I/O facilities to communicate with the
operator's console and peripheral devices, including the disk subsystem.  The
I/O system is accessed by passing a function number and an information address
to CP/M through the FDOS entry point at BOOT+0005H.  In the case of a disk
read, for example, the transient program sends the number corresponding to a
disk read, along with the address of an FCB to the CP/M FDOS.  The FDOS, in
turn, performs the operation and returns with either a disk read completion
indication or an error number indicating that the disk read was unsuccessful.
.sp 2
.tc    5.2  Operating System Call Conventions
.he CP/M Operating System Manual                5.2  Call Conventions
.sh
5.2  Operating System Call Conventions
.qs
.pp
This section provides detailed information for performing direct operating
system calls from user programs.  Many of the functions listed below, however,
are accessed more simply through the I/O macro library provided with the
MAC macro assembler and listed in the Digital Research manual 
entitled, \c
.ul
Programmer's Utilities Guide for the CP/M Family of Operating Systems.
.qu
.pp
CP/M facilities that are available for access by transient programs fall into
two general categories:  simple device I/O and disk file I/O.  The simple
device operations are
.sp
.nf
.in 5
.ti -2
o read a console character
.ti -2
o write a console character
.ti -2
o read a sequential character
.ti -2
o write a sequential character
.ti -2
o get or set I/O status
.ti -2
o print console buffer
.ti -2
o interrogate console ready
.sp
The following FDOS operations perform disk I/O:
.sp
.ti -2
o disk system reset
.ti -2
o drive selection
.ti -2
o file creation
.ti -2
o file close
.ti -2
o directory search
.ti -2
o file delete
.ti -2
o file rename
.ti -2
o random or sequential read
.ti -2
o random or sequential write
.ti -2
o interrogate available disks
.ti -2
o interrogate selected disk
.ti -2
o set DMA address
.ti -2
o set/reset file indicators.
.fi
.in 0
.pp
As mentioned above, access to the FDOS functions is accomplished by passing
a function number and information address through the primary point at
location BOOT+0005H.  In general, the function number is passed in register C
with the information address in the double byte pair DE.  Single byte values
are returned in register A, with double byte values returned in HL, a zero
value is returned when the function number is out of range.  For reasons of
compatibility, register A = L and register B = H upon return in all cases.
Note that the register passing conventions of CP/M agree with
those of the Intel PL/M systems programming language.  CP/M functions and
their numbers are listed below.
.bp
.nf
.in 5
 O  System Reset             19  Delete File
 1  Console Input            20  Read Sequential
 2  Console Output           21  Write Sequential
 3  Reader Input             22  Make File
 4  Punch Output             23  Rename File
 5  List Output              24  Return Login Vector
 6  Direct Console I/O       25  Return Current Disk
 7  Get I/O Byte             26  Set DMA Address
 8  Set I/O Byte             27  Get Addr(Alloc)
 9  Print String             28  Write Protect Disk
10  Read Console Buffer      29  Get R/0 Vector
11  Get Console Status       30  Set File Attributes
12  Return Version Number    31  Get Addr(Disk Parms)
13  Reset Disk System        32  Set/Get User Code
14  Select Disk              33  Read Random
15  Open File                34  Write Random
16  Close File               35  Compute File Size
17  Search for First         36  Set Random Record
18  Search for Next          37  Reset Drive
                             40  Write Random with Zero Fill
.fi
.in 0
.sp
.pp
Functions 28 and 32 should be avoided in application programs to
maintain upward compatibility with CP/M.
.pp
Upon entry to a transient program, the CCP leaves the stack 
pointer set to an eight-level stack area with the CCP return 
address pushed onto the stack, leaving seven levels before 
overflow occurs.  Although this stack is usually not used by a 
transient program (most transients return to the CCP 
through a jump to location 0000H) it is large enough to 
make CP/M system calls because the FDOS switches to a local stack 
at system entry.  For example, the assembly-language program segment below 
reads characters continuously until an asterisk is 
encountered, at which time control returns to the CCP, assuming a 
standard CP/M system with BOOT = 0000H.
.sp 2
.nf
.in 8
BDOS      EQU        0005H      ;STANDARD CP/M ENTRY
CONIN     EQU        1          ;CONSOLE INPUT FUNCTION
;
          ORG        0100H      ;BASE OF TPA
NEXTC:    MVI        C,CONIN    ;READ NEXT CHARACTER
          CALL       BDOS       ;RETURN CHARACTER IN <A>
          CPI        '*'        ;END OF PROCESSING?
          JNZ        NEXTC      ;LOOP IF NOT
          RET                   ;RETURN TO CCP
          END
.fi
.in 0
.sp
.pp
CP/M implements a named file structure on each disk, providing a 
logical organization that allows any particular file to contain 
any number of records from completely empty to the full capacity 
of the drive.  Each drive is logically distinct with a disk 
directory and file data area.  The disk filenames are in three 
parts: the drive select code, the filename (consisting of one to 
eight nonblank characters), and the filetype (consisting of zero 
to three nonblank characters).  The filetype names the generic 
category of a particular file, while the filename distinguishes 
individual files in each category.  The filetypes listed in Table 5-1 
name a few generic categories that have been established, 
although they are somewhat arbitrary.
.sp 2
.sh
                   Table 5-1.  CP/M Filetypes
.qs
.sp
.nf
                Filetype           Meaning
.sp
.in 30
.ti -11
ASM        Assembler Source
.ti -11
PRN        Printer Listing
.ti -11
HEX        Hex Machine Code
.ti -11
BAS        Basic Source File
.ti -11
INT        Intermediate Code
.ti -11
COM        Command File
.ti -11
PLI        PL/I Source File
.ti -11
REL        Relocatable Module
.ti -11
TEX        TEX Formatter Source
.ti -11
BAK        ED Source Backup
.ti -11
SYM        SID Symbol File
.ti -11
$$$        Temporary File
.fi
.in 0
.sp
.pp
Source files are treated as a sequence of ASCII characters, where 
each line of the source file is followed by a carriage return, and 
line-feed sequence (0DH followed by 0AH).  Thus, one 128-byte CP/M 
record can contain several lines of source text.  The end of an 
ASCII file is denoted by a CTRL-Z character (1AH) or a real 
end-of-file returned by the CP/M read operation.  CTRL-Z characters embedded
within machine code files (for example, COM files) are ignored and 
the end-of-file condition returned by CP/M is used to terminate 
read operations.
.pp
Files in CP/M can be thought of as a sequence of up to 65536 
records of 128 bytes each, numbered from 0 through 65535, thus 
allowing a maximum of 8 megabytes per file.  Note, however, 
that although the records may be considered logically 
contiguous, they may not be physically contiguous in the disk 
data area.  Internally, all files are divided into 16K byte 
segments called logical extents, so that counters are easily 
maintained as 8-bit values.  The division into extents is 
discussed in the paragraphs that follow:  however, they are not 
particularly significant for the programmer, because each extent is 
automatically accessed in both sequential and random access  
modes.
.pp
In the file operations starting with Function 15, DE 
usually addresses a FCB.  Transient programs 
often use the default FCB area reserved by CP/M at 
location BOOT+005CH (normally 005CH) for simple file operations.  
The basic unit of file information is a 128-byte record used for 
all file operations.  Thus, a default location for disk I/O is 
provided by CP/M at location BOOT+0080H (normally 0080H) which 
is the initial default DMA address.  See Function 26.
.pp
All directory operations take place in a reserved area that does not 
affect write buffers as was the case in release 1, with the 
exception of Search First and Search Next, where compatibility is 
required.
.pp
The FCB data area consists of a sequence of 33 bytes for 
sequential access and a series of 36 bytes in the case when the 
file is accessed randomly.  The default FCB, normally located at 
005CH, can be used for random access files, because the three bytes 
starting at BOOT+007DH are available for this purpose.  Figure 5-2 shows
the FCB format with the following fields.
.sp 3
.nf
    dr f1 f2 / / f8 t1 t2 t3 ex s1 s2 rc d0 / / dn cr r0 r1 r2
    00 01 02 ... 08 09 10 11 12 13 14 15 16 ... 31 32 33 34 35
.fi
.sp 2
.sh
             Figure 5-2.  File Control Block Format
.sp 3
The following table lists and describes each of the fields in the File Control
Block figure.
.sp 2
.sh
              Table 5-2.  File Control Block Fields
.nf
.sp
       Field                        Definition
.sp
     dr                   drive code (0-16)
                          0 = use default drive for file
                          1 = auto disk select drive A,
                          2 = auto disk select drive B,
                          .
                          .
                          .
                          16= auto disk select drive P.
.sp
     f1...f8              contain the filename in ASCII
                          upper-case, with high bit = 0
.sp
     t1, t2, t3           contain the filetype in ASCII
                          upper-case, with high bit = 0
                          t1', t2', and t3' denote the
                          bit of these positions,
                          t1' = 1 =>Read-Only file,
                          t2' = 1 =>SYS file, no DIR list
.sp
     ex                   contains the current extent
                          number, normally set to 00 by
                          the user, but in range 0-31
                          during file I/O
.bp
.sh
                     Table 5-2.  (continued)
.qs
.sp
       Field                        Definition
.sp
     s1                   reserved for internal system use
.sp
     s2                   reserved for internal system use,
                          set to zero on call to OPEN, MAKE,
                          SEARCH
.sp
     rc                   record count for extent ex;
                          takes on values from 0-127
.sp
     d0...dn              filled in by CP/M; reserved for
                          system use
.sp
     cr                   current record to read or write in
                          a sequential file operation;
                          normally set to zero by user
.sp
     r0, r1, r2           optional random record number in
                          the range 0-65535, with overflow
                          to r2, r0, r1 constitute a 16-bit
                          value with low byte r0, and high
                          byte r1
.fi
.sp
.pp
Each file being accessed through CP/M must have a corresponding 
FCB, which provides the name and allocation information for all 
subsequent file operations.  When accessing files, it is the 
programmer's responsibility to fill the lower 16 bytes of the FCB 
and initialize the cr field.  Normally, bytes 1 through 11 are 
set to the ASCII character values for the filename and filetype,
while all other fields are zero.
.pp
FCBs are stored in a directory area of the disk, and are brought 
into central memory before the programmer proceeds with file 
operations (see the OPEN and MAKE functions).  The memory copy of 
the FCB is updated as file operations take place and later 
recorded permanently on disk at the termination of the file 
operation, (see the CLOSE command).
.pp
The CCP constructs the first 16 bytes of two optional FCBs for a 
transient by scanning the remainder of the line following the 
transient name, denoted by file1 and file2 in the prototype 
command line described above, with unspecified fields set to 
ASCII blanks.  The first FCB is constructed at location 
BOOT+005CH and can be used as is for subsequent file operations.  
The second FCB occupies the d0...dn portion of the first FCB and 
must be moved to another area of memory before use.  If, for 
example, the following command line is typed:
.sp
.ti 8
PROGNAME B:X.ZOT Y.ZAP
.bp
the file PROGNAME.COM is loaded into the TPA, and the default FCB 
at BOOT+005CH is initialized to drive code 2, filename X, and 
filetype ZOT.  The second drive code takes the default value 0, 
which is placed at BOOT-006CH, with the filename Y placed into 
location BOOT+006DH and filetype ZAP located 8 bytes later at 
BOOT+0075H.  All remaining fields through cr are set to zero.  
Note again that it is the programmer's 
responsibility to move this second filename and filetype to another 
area, usually a separate file control block, before opening the 
file that begins at BOOT+005CH, because the open operation 
overwrites the second name and type.
.pp
If no filenames are specified in the original command, the 
fields beginning at BOOT+005DH and BOOT+006DH contain blanks.  In 
all cases, the CCP translates lower-case alphabetics to upper-case
to be consistent with the CP/M file naming conventions.
.pp
As an added convenience, the default buffer area at location 
BOOT+0080H is initialized to the command line tail typed by the 
operator following the program name.  The first position contains 
the number of characters, with the characters themselves 
following the character count.  Given the above command line, the 
area beginning at BOOT+0080H is initialized as follows:
.sp 2
.nf
.in 5
BOOT+0080H:
.sp
+00 +01 +02 +03 +04 +05 +06 +07 +08 +09 +A  +B  +C  +D  +E
E   ''  'B' ':' 'X' '.' 'Z' 'O' 'T' ''  'Y' '.' 'Z' 'A' 'P'
.fi
.in 0
.sp 2
where the characters are translated to upper-case ASCII with 
uninitialized memory following the last valid character.  Again, 
it is the responsibility of the programmer to extract the 
information from this buffer before any file operations are 
performed, unless the default DMA address is explicitly changed.
.pp
Individual functions are described in detail in the pages that 
follow.
.bp
.sp 4
.nf
                    FUNCTION 0:  SYSTEM RESET
.sp
                    Entry Parameters:
                           Register C:  00H
.fi
.sp 2
.pp
The System Reset function returns control to the CP/M operating 
system at the CCP level.  The CCP reinitializes the disk 
subsystem by selecting and logging-in disk drive A.  This 
function has exactly the same effect as a jump to location BOOT.
.sp 6
.nf
                   FUNCTION 1:  CONSOLE INPUT
.sp
                Entry Parameters:
                      Register C:  01H
.sp
                Returned Value:
                      Register A:  ASCII Character
.fi
.sp 2
.pp
The Console Input function reads the next console character to 
register A.  Graphic characters, along with carriage return, line-feed,
and back space (CTRL-H) are echoed to the console.  Tab 
characters, CTRL-I, move the cursor to the next tab stop.  A check 
is made for start/stop scroll, CTRL-S, and start/stop printer echo, 
CTRL-P.  The FDOS does not return to the calling program until a 
character has been typed, thus suspending execution if a 
character is not ready.
.bp
.sp 4
.nf
                   FUNCTION 2:  CONSOLE OUTPUT
.sp
                Entry Parameters
                      Register C:  02H
                      Register E:  ASCII Character
.fi
.sp 2
.pp
The ASCII character from register E is sent to the console 
device.  As in Function 1, tabs are expanded and checks are made 
for start/stop scroll and printer echo.
.sp 6
.nf
                    FUNCTION 3:  READER INPUT
.sp
                Entry Parameters:
                      Register C:  03H
.sp
                Returned Value:
                      Register A:  ASCII Character
.fi
.sp 2
.pp
The Reader Input function reads the next character from the 
logical reader into register A.  See the IOBYTE definition in 
Chapter 6.  Control does not return until the character has been 
read.
.bp
.sp 4
.nf
                    FUNCTION 4:  PUNCH OUTPUT
.sp
                 Entry Parameters:
                       Register C:  04H
                       register E:  ASCII Character
.fi
.sp 2
.pp
The Punch Output function sends the character from register E to 
the logical punch device.
.sp 6
.nf
                    FUNCTION 5:  LIST OUTPUT
.sp
                Entry Parameters:
                      Register C:  05H
                      Register E:  ASCII Character
.fi
.sp 2
.pp
The List Output function sends the ASCII character in register E 
to the logical listing device.
.bp
.sp 4
.nf
                 FUNCTION 6:  DIRECT CONSOLE I/O
.sp
                Entry Parameters:
                      Register C:  06H
                      Register E:  0FFH (input) or
                                   char (output)
.sp
                Returned Value:
                      Register A:  char or status
.fi
.sp 2
.pp
Direct Console I/O is supported under CP/M for those specialized 
applications where basic console input and output are required.  
Use of this function should, in general, be avoided since it 
bypasses all of the CP/M normal control character functions (for example, 
CTRL-S and CTRL-P).  Programs that perform direct I/O 
through the BIOS under previous releases of CP/M, however, should 
be changed to use direct I/O under BDOS so that they can be fully 
supported under future releases of MP/M \ and CP/M.
.pp
Upon entry to Function 6, register E either contains hexadecimal 
FF, denoting a console input request, or an ASCII character.  If 
the input value is FF, Function 6 returns A = 00 if no character 
is ready, otherwise A contains the next console input character.
.pp
If the input value in E is not FF, Function 6 assumes that E 
contains a valid ASCII character that is sent to the console.
.pp
Function 6 must not be used in conjunction with other console I/O 
functions.
.sp 6
.nf
                    FUNCTION 7:  GET I/O BYTE
.sp
                Entry Parameters:
                      Register C:  07H
.sp
                Returned Value:
                      Register A:  I/O Byte Value
.fi
.sp 2
.pp
The Get I/O Byte function returns the current value of IOBYTE in 
register A.  See Chapter 6 for IOBYTE definition.
.bp
.sp 4
.nf
                    FUNCTION 8:  SET I/O BYTE
.sp
                Entry Parameters:
                      Register C:  08H
                      Register E:  I/O Byte Value
.fi
.sp 2
.pp
The SET I/O Byte function changes the IOBYTE value to that given 
in register E.
.sp 6
.nf
                    FUNCTION 9:  PRINT STRING
.sp
               Entry Parameters:
                     Register C:  09H
                   Registers DE:  String Address
.fi
.sp 2
.pp
The Print String function sends the character string stored in 
memory at the location given by DE to the console device, until a 
$ is encountered in the string.  Tabs are expanded as in Function 
2, and checks are made for start/stop scroll and printer echo.
.nx fiveb.tex



