
.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                4-%
.pc 1
.tc 4  CP/M Dynamic Debugging Tool
.nf
.sh
                            Section 4
.sp
.sh
                   CP/M Dynamic Debugging Tool
.qs
.fi
.sp 3
.tc    4.1  Introduction
.he CP/M Operating System Manual                    4.1  Introduction
.sh
4.1  Introduction
.pp 5
The DDT program allows dynamic interactive testing and debugging of
programs generated in the CP/M environment.  Invoke the debugger with
a command of one of the following forms:
.sp
.in 8
.nf
DDT
DDT filename.HEX
DDT filename.COM
.fi
.in 0
.sp
where filename is the name of the program to be loaded and 
tested.  In both cases, the DDT program is brought into main 
memory in place of the Console Command Processor (CCP) and resides 
directly below the Basic Disk Operating System (BDOS)
portion of CP/M.  Refer to Section 5 for standard memory
organization.  The BDOS
starting address, located in the address field of the 
JMP instruction at location 5H, is altered to reflect the 
reduced Transient Program Area (TPA) size.
.pp
The second and third forms of the DDT command perform the same 
actions as the first, except there is a subsequent automatic load 
of the specified HEX or COM file.  The action is identical to the 
following sequence of commands:
.sp
.in 8
.nf
DDT
Ifilename.HEX or Ifilename.COM
R
.fi
.in 0
.sp
where the I and R commands set up and read the specified program 
to test.  See the explanation of the I and R 
commands below for exact details.
.pp
Upon initiation, DDT prints a sign-on message in the form:
.sp
.ti 8
DDT VER m.m
.sp
where m.m is the revision number.
.pp
Following the sign-on message, DDT prompts you with the 
hyphen character, -, and waits for input commands from the 
console.  You can type any of several single-character commands, 
followed by a carriage return to execute the command.  Each 
line of input can be line-edited using the following standard 
CP/M controls:
.bp
.ce
.sh
Table 4-1.  Line-editing Controls
.ll 60
.sp
.in 5
.nf
Control                       Result
.sp
rubout      removes the last character typed

CTRL-U      removes the entire line, ready for retyping

CTRL-C      reboots system
.fi
.in 0
.ll 65
.sp
.pp
Any command can be up to 32 characters in length.  An automatic 
carriage return is inserted as character 33, where the 
first character determines the command type.  Table 4-2 describes DDT
commands.
.sp 2
.sh
                    Table 4-2.  DDT Commands
.sp
.nf
    Command                    Result
   Character
.fi
.sp
.ll 57
.in 16
.ti -9
A        enters assembly-language mnemonics with operands.
.sp
.ti -9
D        displays memory in hexadecimal and ASCII.
.sp
.ti -9
F        fills memory with constant data.
.sp
.ti -9
G        begins execution with optional breakpoints.
.sp
.ti -9
I        sets up a standard input File Control Block.
.sp
.ti -9
L        lists memory using assembler mnemonics.
.sp
.ti -9
M        moves a memory segment from source to destination.
.sp
.ti -9
R        reads a program for subsequent testing.
.sp
.ti -9
S        substitutes memory values.
.sp
.ti -9
T        traces program execution.
.sp
.ti -9
U        untraced program monitoring.
.sp
.ti -9
X        examines and optionally alters the CPU state.
.in 0
.ll 65
.mb 4
.fm 1
.sp 2
The command character, in some cases, is followed by zero, one, 
two, or three hexadecimal values, which are separated by commas 
or single blank characters.  All DDT numeric output is in 
hexadecimal form.  The commands are not execution until the 
carriage return is typed at the end of the command.
.pp
At any point in the debug run, you can stop execution of 
DDT by using either a CTRL-C or G0 (jump to location 0000H) and 
save the current memory image by using a SAVE command of the form:
.sp
.ti 8
SAVE n filename. COM
.sp
where n is the number of pages (256 byte blocks) to be saved on 
disk.  The number of blocks is determined by taking the high-order
byte of the address in the TPA and converting this number to 
decimal.  For example, if the highest address in the TPA is 134H, 
the number of pages is 12H or 18 in decimal.  You could type a 
CTRL-C during the debug run, returning to the CCP level, followed 
by
.mb 6
.fm 2
.sp
.ti 8
SAVE 18 X.COM
.sp
The memory image is saved as X.COM on the disk and can be 
directly executed by typing the name X.  If further testing is 
required, the memory image can be recalled by typing
.sp
.ti 8
DDT X.COM
.sp
which reloads the previously saved program from location 100H 
through page 18, 23FFH.  The CPU state is not a part of the COM 
file; thus, the program must be restarted from the beginning to 
test it properly.
.sp 2
.tc    4.2  DDT Commands
.he CP/M Operating System Manual                    4.2  DDT Commands
.sh
4.2  DDT Commands
.pp
The individual commands are detailed below.  In each case, the 
operator must wait for the hyphen prompt character before entering 
the command.  If control is passed to a program under test, and 
the program has not reached a breakpoint, control can be returned 
to DDT by executing a RST 7 from the front panel.  In the 
explanation of each command, the command letter is shown in some 
cases with numbers separated by commas, the the numbers are 
represented by lower-case letters.  These numbers are always 
assumed to be in a hexadecimal radix and from one to four digits 
in length.  Longer numbers are automatically truncated on the 
right.
.pp
Many of the commands operate upon a CPU state that corresponds 
to the program under test.  The CPU state holds the registers of 
the program being debugged and initially contains zeros for all 
registers and flags except for the program counter, P, and stack 
pointer, S, which default to 100H.  The program counter is 
subsequently set to the starting address given in the last record 
of a HEX file if a file of this form is loaded, see the I and R 
commands.
.sp 2
.tc         4.2.1  The A (Assembly) Command
.sh
4.2.1  The A (Assembly) Command
.pp
DDT allows in-line assembly language to be inserted into the 
current memory image using the A command, which takes the form:
.sp
.ti 8
As
.sp
where s is the hexadecimal starting address for the in-line 
assembly.  DDT prompts the console with the address of the next 
instruction to fill and reads the console, looking for assembly-language
mnemonics followed by register references and operands in
absolute hexadecimal form.  See the \c
.ul
Intel 8080 Assembly Language Reference Card \c
.qu
for a list of mnemonics.  Each 
successive load address is printed before reading the console.  
The A command terminates when the first empty line is input from 
the console.
.pp
Upon completion of assembly language input, you can 
review the memory segment using the DDT disassembler (see the L 
command).
.pp
Note that the assembler/disassembler portion of 
DDT can be overlaid by the transient program being tested, in 
which case the DDT program responds with an error condition when 
the A and L commands are used.
.sp 2
.tc         4.2.2  The D (Display) Command
.sh
4.2.2  The D (Display) Command
.pp
The D command allows you to view the contents of memory 
in hexadecimal and ASCII formats.  The D command takes the forms:
.sp
.in 8
.nf
D
Ds
Ds,f
.fi
.in 0
.pp
In the first form, memory is displayed from the current display 
address, initially 100H, and continues for 16 display lines.  
Each display line takes the followng form:
.sp
.nf
aaaa bb bb bb bb bb bb bb bb bb bb bb bb bb bb bb bb cccccccccccccccc
.fi
.sp
where aaaa is the display address in hexadecimal and bb 
represents data present in memory starting at aaaa.  The ASCII 
characters starting at aaaa are to the right (represented by the 
sequence of character c) where nongraphic characters are printed as a 
period.  You should note that both upper- and lower-case 
alphabetics are displayed, and will appear as upper-case symbols 
on a console device that supports only upper-case.  Each display 
line gives the values of 16 bytes of data, with the first line 
truncated so that the next line begins at an address that is a 
multiple of 16.
.pp
The second form of the D command is similar to the first, except 
that the display address is first set to address s.
.pp
The third form causes the display to continue from address s 
through address f.  In all cases, the display address is set to 
the first address not displayed in this command, so that a 
continuing display can be accomplished by issuing successive D 
commands with no explicit addresses.
.pp
Excessively long displays can be aborted by pressing the return key.
.sp 2
.tc         4.2.3  The F (Fill) Command
.sh
4.2.3  The F (Fill) Command
.pp
The F command takes the form:
.sp
.ti 8
Fs,f,c,
.sp
where s is the starting address, f is the final address, and c is 
a hexadecimal byte constant.  DDT stores the constant c at 
address s, increments the value of s and test against f.  If s 
exceeds f, the operation terminates, otherwise the operation is 
repeated.  Thus, the fill command can be used to set a memory 
block to a specific constant value.
.sp 2
.tc         4.2.4  The G (Go) Command
.sh
4.2.4  The G (Go) Command
.pp
A program is executed using the G command, with up to two 
optional breakpoint addresses.  The G command takes the forms:
.sp 2
.in 8
.nf
G
Gs
Gs,b
Gs,b,c
G,b
G,b,c
.fi
.in 0
.sp
.pp
The first form executes the program at the current value of the 
program counter in the current machine state, with no breakpoints 
set.  The only way to regain control in DDT is through a RST 7 
execution.  The current program counter can be viewed by typing 
an X or XP command.
.pp
The second form is similar to the first, except that the program 
counter in the current machine state is set to address s before 
execution begins.
.pp
The third form is the same as the second, except that program 
execution stops when address b is encountered (b must be in the 
area of the program under test).  The instruction at location b is 
not executed when the breakpoint is encountered.
.pp
The fourth form is identical to the third, except that two 
breakpoints are specified, one at b and the other at c.  
Encountering either breakpoint causes execution to stop, and both 
breakpoints are cleared.  The last two forms take the program 
counter from the current machine state and set one and two 
breakpoints, respectively.
.pp
Execution continues from the starting address in real-time to the 
next breakpoint.  There is no intervention between the starting 
address and the break address by DDT.  If the program under test 
does not reach a breakpoint, control cannot return to DDT without 
executing a RST 7 instruction.  Upon encountering a breakpoint, 
DDT stops execution and types
.sp
.ti 8
*d
.sp
where d is the stop address.  The machine state can be examined 
at this point using the X (Examine) command.  You must 
specify breakpoints that differ from the program counter address 
at the beginning of the G command.  Thus, if the current program 
counter is 1234H, then the following commands:
.sp
.nf
.in 8
G,1234
G400,400
.fi
.in 0
.sp
both produce an immediate breakpoint without executing any 
instructions.
.sp 2
.tc         4.2.5  The I (Input) Command
.sh
4.2.5  The I (Input) Command
.pp
The I command allows you to insert a filename into the default 
File Control Block (FCB) at 5CH.  The FCB created by CP/M for 
transient programs is placed at this location (see Section 5).  
The default FCB can be used by the program under test as if it 
had been passed by the CP/M Console Processor.  Note that this 
filename is also used by DDT for reading additional HEX and COM 
files.  The I command takes the forms:
.sp
.nf
.in 8
Ifilename
Ifilename.typ
.fi
.in 0
.pp
If the second form is used and the filetype is either HEX or COM, 
subsequent R commands can be used to read the pure binary or hex 
format machine code.  Section 4.2.8 gives further details.
.sp 2
.tc         4.2.6  The L (List) Command
.sh
4.2.6  The L (List) Command
.pp
The L command is used to list assembly-language mnemonics in a 
particular program region.  The L command takes the forms:
.sp
.in 8
.nf
L
Ls
Ls,f
.fi
.in 0
.pp
The first form lists twelve lines of disassembled machine code 
from the current list address.  The second form sets the list 
address to s and then lists twelve lines of code.  The last form 
lists disassembled code from s through address f.  In all three 
cases, the list address is set to the next unlisted location in 
preparation for a subsequent L command.  Upon encountering an 
execution breakpoint, the list address is set to the current 
value of the program counter (G and T commands).  Again, long 
typeouts can be aborted by pressing RETURN during the list 
process.
.sp 2
.tc         4.2.7  The M (Move) Command
.sh
4.2.7  The M (Move) Command
.pp
The M command allows block movement of program or data areas from 
one location to another in memory.  The M command takes the form:
.sp
.ti 8
Ms,f,d
.sp
where s is the start address of the move, f is the final address, 
and d is the destination address.  Data is first removed from s 
to d, and both addresses are incremented.  If s exceeds f, the 
move operation stops; otherwise, the move operation is repeated.
.sp 2
.tc         4.2.8  The R (Read) Command
.sh
4.2.8  The R (Read) Command
.pp
The R command is used in conjunction with the I command to read 
COM and HEX files from the disk into the transient program 
area in preparation for the debug run.  The R command takes the forms:
.sp
.in 8
.nf
R
RB
.fi
.in 0
.sp
where b is an optional bias address that is added to each program 
or data address as it is loaded.  The load operation must not 
overwrite any of the system parameters from 000H through 0FFH
(that is, the first page of memory).  If b is omitted, then 
b=0000 is assumed.  The R command requires a previous I command, 
specifying the name of a HEX or COM file.  The load address for 
each record is obtained from each individual HEX record, while an 
assumed load address of 100H is used for COM files.  Note that 
any number of R commands can be issued following the I command to 
reread the program under test, assuming the tested program does 
not destroy the default area at 5CH.  Any file specified with the 
filetype COM is assumed to contain machine code in pure binary 
form (created with the LOAD or SAVE command), and all others are 
assumed to contain machine code in Intel hex format (produced, 
for example, with the ASM command).
.pp
Recall that the command,
.sp
.ti 8
DDT filename.filetype
.sp
which initiates the DDT program, equals to the following commands:
.sp
.in 8
.nf
DDT
-Ifilename.filetype
-R
.fi
.in 0
.bp
.pp
Whenever the R command is issued, DDT responds with either the 
error indicator ? (file cannot be opened, or a checksum error 
occurred in a HEX file) or with a load message.  The load message
takes the form:
.sp
.in 8
.nf
NEXT PC
nnnn pppp
.fi
.in 0
.sp
where nnnn is the next address following the loaded program and 
pppp is the assumed program counter (100H for COM files, or 
taken from the last record if a HEX file is specified).
.sp 2
.tc         4.2.9  The S (Set) Command
.sh
4.2.9  The S (Set) Command
.pp
The S command allows memory locations to be examined and 
optionally altered.  The S command takes the form:
.sp
.ti 8
Ss
.sp
where s is the hexadecimal starting address for examination and 
alteration of memory.  DDT responds with a numeric prompt, giving 
the memory location, along with the data currently held in 
memory.  If you type a carriage return, the data is 
not altered.  If a byte value is typed, the value is stored at 
the prompted address.  In either case, DDT continues to prompt 
with successive addresses and values until you type either a period 
or an invalid input value is detected.
.sp 2
.tc         4.2.10 The T (Trace) Command
.sh
4.2.10  The T (Trace) Command
.pp
The T command allows selective tracing of program execution for 1 
to 65535 program steps.  The T command takes the forms:
.sp
.in 8
.nf
T
Tn
.fi
.in 0
.mb 4
.fm 1
.pp
In the first form, the CPU state is displayed and the next 
program step is executed.  The program terminates immediately, 
with the termination address displayed as
.sp
.ti 8
*hhhh
.sp
where hhhh is the next address to execute.  The display address 
(used in the D command) is set to the value of H and L, and the 
list address (used in the L command) is set to hhhh.  The CPU 
state at program termination can then be examined using the X 
command.
.pp
The second form of the T command is similar to the first, except 
that execution is traced for n steps (n is a hexadecimal value) 
before a program breakpoint occurs.  A breakpoint can be forced 
in the trace mode by typing a rubout character.  The CPU state is 
displayed before each program step is taken in trace mode.  The 
format of the display is the same as described in the X command.
.mb 6
.fm 2
.pp
You should note that program tracing is discontinued at the 
CP/M interface and resumes after return from CP/M to the program 
under test.  Thus, CP/M functions that access I/O devices, such 
as the disk drive, run in real-time, avoiding I/O timing 
problems.  Programs running in trace mode execute approximately 
500 times slower than real-time because DDT gets control after each 
user instruction is executed.  Interrupt processing routines can 
be traced, but commands that use the breakpoint facility (G, T, 
and U) accomplish the break using an RST 7 instruction, which 
means that the tested program cannot use this interrupt location.  
Further, the trace mode always runs the tested program with 
interrupts enabled, which may cause problems if asynchronous 
interrupts are received during tracing.
.pp
To get control back to DDT during trace, press RETURN rather than 
executing an RST 7.  This ensures that the trace for current 
instruction is completed before interruption.
.sp 2
.tc         4.2.11 The U (Untrace) Command
.sh
4.2.11  The U (Untrace) Command
.pp
The U command is identical to the T command, except that 
intermediate program steps are not displayed.  The untrace mode 
allows from 1 to 65535, (0FFFFH) steps to be executed in monitored 
mode and is used principally to retain control of an executing 
program while it reaches steady state conditions.  All conditions 
of the T command apply to the U command.
.sp 2
.tc         4.2.12 The X (Examine) Command
.sh
4.2.12  The X (Examine) Command
.pp
The X command allows selective display and alteration of the 
current CPU state for the program under test.  The X command 
takes the forms:
.sp
.in 8
.nf
X
Xr
.fi
.in 0
.sp
where r is one of the 8080 CPU registers listed in the following table.
.sp 2
.sh
                    Table 4-3.  CPU Registers
.sp
.nf
            Register        Meaning           Value
.sp
               C        Carry flag            (0/1)
               Z        Zero flag             (0/1)
               M        Minus flag            (0/1)
               E        Even parity flag      (0/1)
               I        Interdigit carry      (0/1)
               A        Accumulator           (0-FF)
               B        BC register pair      (0-FFFF)
               D        DE register pair      (0-FFFF)
.bp
.sh
                     Table 4-3.  (continued)
.sp
.nf
            Register        Meaning           Value
.sp
               H        HL register pair      (0-FFFF)
               S        Stack pointer         (0-FFFF)
               P        Program counter       (0-FFFF)
.fi
.sp 2
In the first case, the CPU register state is displayed in the 
format:
.sp
.ti 8
CfZfMfEflf A=bb B=dddd D=dddd H=dddd S=dddd P=dddd inst
.sp
where f is a 0 or 1 flag value, bb is a byte value, and dddd is a 
double-byte quantity corresponding to the register pair.  The 
inst field contains the disassembled instruction, that occurs at 
the location addressed by the CPU state's program counter.
.pp
The second form allows display and optional alteration of 
register values, where r is one of the registers given above (C, 
Z, M, E, I, A, B, D, H, S, or P).  In each case, the flag or 
register value is first displayed at the console.  The DDT 
program then accepts input from the console.  If a carriage 
return is typed, the flag or register value is not altered.  If 
a value in the proper range is typed, the flag or register value 
is altered.  You should note that BC, DE, and HL are 
displayed as register pairs.  Thus, you must type the entire 
register pair when B, C, or the BC pair is altered.
.sp 2
.tc    4.3  Implementation Notes
.he CP/M Operating System Manual            4.3  Implementation Notes
.sh
4.3  Implementation Notes
.pp
The organization of DDT allows certain nonessential portions to 
be overlaid to gain a larger transient program area for debugging 
large programs.  The DDT program consists of two parts:  the DDT 
nucleus and the assembler/disassembler module.  The DDT nucleus 
is loaded over the CCP and, although loaded with the DDT nucleus, 
the assembler/disassembler is overlayable unless used to assemble 
or disassemble.
.pp
In particular, the BDOS address at location 6H (address field of 
the JMP instruction at location 5H) is modified by DDT to address 
the base location of the DDT nucleus, which, in turn, contains a 
JMP instruction to the BDOS.  Thus, programs that use this 
address field to size memory see the logical end of memory at the 
base of the DDT nucleus rather than the base of the BDOS.
.pp
The assembler/disassembler module resides directly below the DDT 
nucleus in the transient program area.  If the A, L, T, or X 
commands are used during the debugging process, the DDT program 
again alters the address field at 6H to include this module, 
further reducing the logical end of memory.  If a program loads 
beyond the beginning of the assembler/disassembler module, the A 
and L commands are lost (their use produces a ? in response)
and the trace and display (T and X) commands list the inst field 
of the display in hexadecimal, rather than as a decoded 
instruction.
.nx fourb



