.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                6-%
.pc 1
.tc 6  CP/M 2 Alteration
.ce
.sh
Section 6
.qs
.sp
.ce
.sh
CP/M 2 Alteration
.qs
.sp 3
.tc    6.1  Introduction
.he CP/M Operating System Manual                    6.1  Introduction
.sh
6.1  Introduction
.qs
.pp
The standard CP/M system assumes operation on an Intel Model 
800 microcomputer development system , but is designed so you can alter a
specific set of subroutines that define the hardware operating 
environment.
.pp
Although standard CP/M 2 is configured for single-density floppy 
disks, field-alteration features allow adaptation to a wide 
variety of disk subsystems from single-drive minidisks to 
high-capacity, hard disk systems.  To simplify the following 
adaptation process, it is assumed that CP/M 2 is first 
configured for single-density floppy disks where minimal editing 
and debugging tools are available.  If an earlier version of CP/M 
is available, the customizing process is eased considerably.  In 
this latter case, you might want to review the system 
generation process and skip to later sections that discuss system 
alteration for nonstandard disk systems.
.pp
To achieve device independence, CP/M is separated into three 
distinct modules:
.sp
.in 5
.ti -2
o BIOS is the Basic I/O System, which is environment dependent.
.sp
.ti -2
o BDOS is the Basic Disk Operating System, which is not dependent upon the
hardware configuration.
.sp
.ti -2
o CCP is the Console Command Processor, which uses the BDOS.
.fi
.in 0
.pp
Of these modules, only the BIOS is dependent upon the particular 
hardware.  You can patch the distribution version 
of CP/M to provide a new BIOS that provides a customized 
interface between the remaining CP/M modules and the 
hardware system.  This document provides a step-by-step 
procedure for patching a new BIOS into CP/M.
.mb 4
.fm 1
.pp
All disk-dependent portions of CP/M 2 are placed into a BIOS, a 
resident disk parameter block, which is either hand coded or 
produced automatically using the disk definition macro library 
provided with CP/M 2.  The end user need only specify the maximum 
number of active disks, the starting and ending sector numbers, 
the data allocation size, the maximum extent of the logical disk, 
directory size information, and reserved track values.
The macros use this information to generate 
the appropriate tables and table references for use during CP/M 2 
operation.  Deblocking information is provided, which aids in 
assembly or disassembly of sector sizes that are multiples of the 
fundamental 128-byte data unit, and the system alteration manual 
includes general purpose subroutines that use the deblocking 
information to take advantage of larger sector sizes.  Use of 
these subroutines, together with the table-drive data access 
algorithms, makes CP/M 2 a universal data management system.
.pp
File expansion is achieved by providing up to 512 logical file 
extents, where each logical extent contains 16K bytes of data.  
CP/M 2 is structured, however, so that as much as 128K bytes of 
data are addressed by a single physical extent, corresponding to a 
single directory entry, maintaining compatibility with previous 
versions while taking advantage of directory space.
.pp
If CP/M is being tailored to a computer system for the first 
time, the new BIOS requires some simple software development and 
testing.  The standard BIOS is listed in Appendix A and can be 
used as a model for the customized package.  A skeletal version 
of the BIOS given in Appendix B can serve as the basis for a 
modified BIOS.
.mb 6
.fm 2
.pp
In addition to the BIOS, you must write a simple memory 
loader, called GETSYS, which brings the operating system into 
memory.  To patch the new BIOS into CP/M, you must write the 
reverse of GETSYS, called PUTSYS, which places an altered version 
of CP/M back onto the disk.  PUTSYS can be derived from GETSYS by 
changing the disk read commands into disk write commands.  Sample 
skeletal GETSYS and PUTSYS programs are described in Section 6.4 
and listed in Appendix C.
.pp
To make the CP/M system load automatically, you must also 
supply a cold start loader, similar to the one provided with 
CP/M, listed in Appendixes A and D.  A skeletal form of a cold 
start loader is given in Appendix E, which serves as a model for 
the loader.
.mb 4
.fm 1
.sp 2
.tc    6.2  First-level System Regeneration
.he CP/M Operating System Manual        6.2  First-level Regeneration
.sh
6.2  First-level System Regeneration
.qs
.pp
The procedure to patch the CP/M system is given below.  Address 
references in each step are shown with H denoting the 
hexadecimal radix, and are given for a 20K CP/M system.  For 
larger CP/M systems, a bias is added to each address that is 
shown with a +b following it, where b is equal to the memory 
size-20K.  Values for b in various standard memory sizes are listed in
Table 6-1.
.sp 2
.sh
             Table 6-1.  Standard Memory Size Values

.nf
          Memory Size                Value
.fi
.sp
.in 13
24K:         b = 24K - 20K =  4K = 1000H
.sp
32K:         b = 32K - 20K = 12K = 3000H
.sp
40K:         b = 40K - 20K = 20K = 5000H
.sp
48K:         b = 48K - 20K = 28K = 7000H
.sp
56K:         b = 56K - 20K = 36K = 9000H
.sp
62K:         b = 62K - 20K = 42K = A800H
.sp
64K:         b = 64K - 20K = 44K = B000H
.fi
.in 0
.pp
Note that the standard distribution version of CP/M is set for
operation within a 20K CP/M system. Therefore, you must first bring up
the 20K CP/M system, then configure it for actual 
memory size (see Section 6.3).
.pp
Follow these steps to patch your CP/M system:
.sp 2
.in 8
.ti -3
1) Read Section 6.4 and write a GETSYS program that reads the 
first two tracks of a disk into memory.  The program from the 
disk must be loaded starting at location 3380H.  GETSYS is coded 
to start at location 100H (base of the TPA) as shown in Appendix 
C.
.mb 6
.fm 2
.sp
.ti -3
2) Test the GETSYS program by reading a blank disk into memory, 
and check to see that the data has been read properly and that 
the disk has not been altered in any way by the GETSYS program.
.sp
.ti -3
3) Run the GETSYS program using an initialized CP/M disk to see 
if GETSYS loads CP/M starting at 3380H (the operating system 
actually starts 128 bytes later at 3400H).
.sp
.ti -3
4) Read Section 6.4 and write the PUTSYS program.  This writes 
memory starting at 3380H back onto the first two tracks of the 
disk.  The PUTSYS program should be located at 200H, as shown in 
Appendix C.
.sp
.ti -3
5) Test the PUTSYS program using a blank, uninitialized disk by 
writing a portion of memory to the first two tracks; clear memory 
and read it back using GETSYS.  Test PUTSYS completely, because 
this program will be used to alter CP/M on disk.
.sp
.ti -3
6) Study Sections 6.5, 6.6, and 6.7 along with the distribution 
version of the BIOS given in Appendix A and write a simple 
version that performs a similar function for the customized 
environment.  Use the program given in Appendix B as a model.  
Call this new BIOS by name CBIOS (customized BIOS).  Implement 
only the primitive disk operations on a single drive and simple 
console input/output functions in this phase.
.sp
.ti -3
7) Test CBIOS completely to ensure that it properly performs 
console character I/O and disk reads and writes.  Be careful to 
ensure that no disk write operations occur during read operations 
and check that the proper track and sectors are addressed on all 
reads and writes.  Failure to make these checks might cause 
destruction of the initialized CP/M system after it is patched.
.mb 4
.fm 1
.sp
.ti -3
8) Referring to Table 6-3 in Section 6.5, note that the BIOS is 
placed between locations 4A00H and 4FFFH.  Read the CP/M system 
using GETSYS and replace the BIOS segment by the CBIOS developed 
in step 6 and tested in step 7.  This replacement is done in 
memory.
.sp
.ti -3
9) Use PUTSYS to place the patched memory image of CP/M onto the 
first two tracks of a blank disk for testing.
.sp
.ti -4
10) Use GETSYS to bring the copied memory image from the test 
disk back into memory at 3380H and check to ensure that it has 
loaded back properly (clear memory, if possible, before the 
load).  Upon successful load, branch to the cold start code at 
location 4A00H.  The cold start routine initializes page 
zero, then jumps to the CCP at location 3400H, which calls the 
BDOS, which calls the CBIOS.  The CCP asks the CBIOS to read 
sixteen sectors on track 2, and CP/M types A>, the system 
prompt.
.mb 6
.fm 2
.sp
If difficulties are encountered, use whatever debug facilities 
are available to trace and breakpoint the CBIOS.
.sp
.ti -4
11) Upon completion of step 10, CP/M has prompted the console for 
a command input.  To test the disk write operation, type
.sp
SAVE 1 X.COM
.sp
All commands must be followed by a carriage return.  CP/M 
responds with another prompt after several disk accesses:
.sp
A>
.sp
If it does not, debug the disk write functions and retry.
.sp
.ti -4
12) Test the directory command by typing
.sp
DIR
.sp
CP/M responds with
.sp
A:X     COM
.sp
.ti -4
13) Test the erase command by typing
.sp
ERA X.COM
.sp
CP/M responds with the A prompt.  This is now an operational 
system that only requires a bootstrap loader to function 
completely.
.sp
.ti -4
14) Write a bootstrap loader that is similar to GETSYS and place 
it on track 0, sector 1, using PUTSYS (again using the test disk, 
not the distribution disk).  See Sections 6.5 and 6.8 for more 
information on the bootstrap operation.
.sp
.ti -4
15) Retest the new test disk with the bootstrap loader installed 
by executing steps 11, 12, and 13.  Upon completion of these 
tests, type a CTRL-C.  The system executes a warm start, which 
reboots the system, and types the A prompt.
.sp
.ti -4
16) At this point, there is probably a good version of the 
customized CP/M system on the test disk.  Use GETSYS to load CP/M 
from the test disk.  Remove the test disk, place the distribution 
disk, or a legal copy, into the drive, and use PUTSYS to 
replace the distribution version with the customized version.  
Do not make this replacement if you are unsure of the patch 
because this step destroys the system that was obtained from 
Digital Research.
.sp
.ti -4
17) Load the modified CP/M system and test it by typing
.sp
DIR
.sp
CP/M responds with a list of files that are provided on the 
initialized disk.  The file DDT.COM is the memory image for the 
debugger.  Note that from now on, you must always reboot the 
CP/M system (CTRL-C is sufficient) when the disk is removed and 
replaced by another disk, unless the new disk is to be Read-Only.
.sp
.ti -4
18) Load and test the debugger by typing
.sp
DDT
.sp
See Chapter 4 for operating procedures.
.sp
.ti -4
19) Before making further CBIOS modifications, practice using the 
editor (see Chapter 2), and assembler (see Chapter 3).  Recode 
and test the GETSYS, PUTSYS, and CBIOS programs using ED, ASM, 
and DDT.  Code and test a COPY program that does a sector-to-sector
copy from one disk to another to obtain back-up copies of 
the original disk.  Read the CP/M Licensing Agreement specifying 
legal responsibilities when copying the CP/M system.  Place the 
following copyright notice:
.sp
.nf
Copyright (c), 1983
 Digital Research
.fi
.sp
on each copy that is made with the COPY program.
.sp
.ti -4
20) Modify the CBIOS to include the extra functions for punches, 
readers, and sign-on messages, and add the facilities for 
additional disk drives, if desired.  These changes can be made 
with the GETSYS and PUTSYS programs or by referring to the 
regeneration process in Section 6.3.
.fi
.in 0
.sp
.pp
You should now have a good copy of the customized CP/M 
system.  Although the CBIOS portion of CP/M belongs to the user, 
the modified version cannot be legally copied.
.pp
It should be noted that the system remains file-compatible with 
all other CP/M systems (assuming media compatibility) which 
allows transfer of nonproprietary software between CP/M users.
.tc    6.3  Second-level System Generation
.bp
.he CP/M Operating System Manual  6.3  Second-level System Generation
.sh
6.3  Second-level System Generation
.qs
.pp
Once the system is running, the next step is to configure CP/M 
for the desired memory size.  Usually, a memory image is first 
produced with the MOVCPM program (system relocator) and then 
placed into a named disk file.  The disk file can then be loaded, 
examined, patched, and replaced using the debugger and the 
system generation program (refer to Chapter 1).
.pp
The CBIOS and BOOT are modified using ED and assembled using ASM, 
producing files called CBIOS.HEX and BOOT.HEX, which contain the 
code for CBIOS and BOOT in Intel hex format.
.pp
To get the memory image of CP/M into the TPA configured for the 
desired memory size, type the command:
.sp
.ti 8
MOVCPM xx*
.sp
where xx is the memory size in decimal K bytes, for example, 32 
for 32K.  The response is as follows:
.sp
.nf
.in 8
CONSTRUCTING xxK CP/M VERS 2.0
.sp
READY FOR "SYSGEN" OR
.sp
"SAVE 34 CPMxx.COM"
.fi
.in 0
.pp
An image of CP/M in the TPA is configured for the requested 
memory size.  The memory image is at location 0900H through 
227FH, that is, the BOOT is at 0900H, the CCP is at 980H, the 
BDOS starts at 1180H, and the BIOS is at 1F80H.  Note that the 
memory image has the standard Model 800 BIOS and BOOT on it.  It is now 
necessary to save the memory image in a file so that you can 
patch the CBIOS and CBOOT into it:
.sp
.ti 8
SAVE 34 CPMxx.COM
.pp
The memory image created by the MOVCPM program is offset by a 
negative bias so that it loads into the free area of the TPA, and 
thus does not interfere with the operation of CP/M in higher 
memory.  This memory image can be subsequently loaded under DDT 
and examined or changed in preparation for a new generation of 
the system.  DDT is loaded with the memory image by typing:
.sp
.ti 8
DDT CPMxx.COM       Loads DDT, then reads the CP/M image.
.sp
DDT should respond with the following:
.sp
.nf
.in 8
NEXT PC
2300 0100
-                                          The DDT prompt
.fi
.in 0
.sp
You can then give the display and disassembly commands to examine
portions of the memory image between 900H and 227FH.
Note, however, that to find any particular address 
within the memory image, you must apply the negative bias to the 
CP/M address to find the actual address.  Track 00, sector 01, is 
loaded to location 900H (the user should find the cold start 
loader at 900H to 97FH); track 00, sector 02, is loaded into 980H 
(this is the base of the CCP); and so on through the entire CP/M 
system load.  In a 20K system, for example, the CCP resides at 
the CP/M address 3400H, but is placed into memory at 980H by the 
SYSGEN program.  Thus, the negative bias, denoted by n, satisfies
.sp
.ti 8
3400H + n = 980H, or n =980H - 3400H
.sp
Assuming two's complement arithmetic, n = D580H, which can be 
checked by
.sp
.ti 8
.nf
3400H + D580H = 10980H = 0980H (ignoring high-order
                                       overflow).
.fi
.pp
Note that for larger systems, n satisfies
.sp
.nf
.in 8
(3400H+b) + n = 980H, or
n = 980H - (3400H + b), or
n = D580H - b
.fi
.in 0
.sp
The value of n for common CP/M systems is given below.
.sp 2
.sh
           Table 6-2.  Common Values for CP/M Systems
.sp
.nf
         Memory Size      BIAS b      Negative Offset n
.sp
.in 13
20K          0000H     D580H - 0000H = D580H
24K          1000H     D580H - 1000H = C580H
32K          3000H     D580H - 3000H = A580H
40K          5000H     D580H - 5000H = 8580H
48K          7000H     D580H - 7000H = 6580H
56K          9000H     D580H - 9000H = 4580H
62K          A800H     D580H - A800H = 2D80H
64K          B000H     D580H - B000H = 2580H
.fi
.in 0
.sp
.pp
If you want to locate the address x within the memory image 
loaded under DDT in a 20K system, first type
.sp
.ti 8
Hx,n            Hexadecimal sum and difference
.sp
and DDT responds with the value of x+n (sum) and x-n 
(difference).  The first number printed by DDT is the actual memory 
address in the image where the data or code is located.  For example,
the following DDT command:
.sp
.ti 8
H3400,D580
.sp
produces 980H as the sum, which is where the CCP 
is located in the memory image under DDT.
.pp
Type the L command to disassemble portions of the 
BIOS located at (4A00H+b)-n, which, when one uses the H command, 
produces an actual address of 1F80H.  The disassembly command 
would thus be as follows:
.sp
.ti 8
L1F80
.sp
It is now necessary to patch in the CBOOT and CBIOS routines.  The BOOT
resides at location 0900H in the memory image.  If the actual 
load address is n, then to calculate the bias (m), 
type the command:
.sp
.ti 8
H900,n          Subtract load address from target address.
.pp
The second number typed by DDT in response to the command is the 
desired bias (m).  For example, if the BOOT executes at 0080H, 
the command
.sp
.ti 8
H900,80
.sp
produces
.sp
.ti 8
0980 0880       Sum and difference in hex.
.sp
Therefore, the bias m would be 0880H.  To read-in the BOOT, give the command:
.sp
.ti 8
ICBOOT.HEX      Input file CBOOT.HEX
.sp
Then
.sp
.ti 8
Rm              Read CBOOT with a bias of m (=900H-n).
.sp
Examine the CBOOT with
.sp
.ti 8
L900
.sp
You are now ready to replace the CBIOS by examining the area at 
1F80H, where the original version of the CBIOS resides, and then 
typing
.sp
.ti 8
ICBIOS.HEX      Ready the hex file for loading.
.pp
Assume that the CBIOS is being integrated into a 20K 
CP/M system and thus originates at location 4A00H.  To locate the 
CBIOS properly in the memory image under DDT, you must apply the 
negative bias n for a 20K system when loading the hex file.  This 
is accomplished by typing
.sp
.ti 8
RD580           Read the file with bias D580H.
.sp
Upon completion of the read, reexamine the area 
where the CBIOS has been loaded (use an L1F80 command) to ensure 
that it is properly loaded.  When you are satisfied that the change has 
been made, return from DDT using a CTRL-C or, G0 command.
.pp
SYSGEN is used to replace the patched memory image back onto a 
disk (you use a test disk until sure of the 
patch) as shown in the following interaction:
.sp 2
.nf
.in 8
SYSGEN                    Start the SYSGEN program.
.sp
SYSGEN VERSION 2.0        Sign-on message from SYSGEN.
.sp
SOURCE DRIVE NAME         Respond with a carriage return
(OR RETURN TO SKIP)       to skip the CP/M read operation
                          because the system is already
                          in memory.
.sp
DESTINATION DRIVE NAME    Respond with B to write the new
(OR RETURN TO REBOOT)     system to the disk in drive B.

.sp
DESTINATION ON B,         Place a scratch disk in drive
THEN TYPE RETURN          B, then press RETURN.
.sp
FUNCTION COMPLETE
DESTINATION DRIVE NAME
(OR RETURN TO REBOOT)
.fi
.in 0
.sp
.pp
Place the scratch disk in drive A, then 
perform a cold start to bring up the newly-configured CP/M 
system.
.pp
The new CP/M system is then tested and the Digital Research 
copyright notice is placed on the disk, as specified in the 
Licensing Agreement:
.sp
.nf
.in 8
Copyright (c), 1979
 Digital Research
.fi
.in 0
.sp 2
.tc    6.4  Sample GETSYS and PUTSYS Programs
.he CP/M Operating System Manual        6.4  Sample GETSYS and PUTSYS
.sh
6.4  Sample GETSYS and PUTSYS Programs
.qs
.pp
The following program provides a framework for the GETSYS and 
PUTSYS programs referenced in Sections 6.1 and 6.2.  To read and
write the specific sectors, you must insert the READSEC and WRITESEC
subroutines.
.bp
.nf
;  GETSYS PROGRAM -- READ TRACKS 0 AND 1 TO MEMORY AT 3380H
;  REGISTER                    USE
.sp
;         A               (SCRATCH REGISTER)
.sp
;         B               TRACK COUNT (0, 1)
.sp
;         C               SECTOR COUNT (1,2,...,26)
.sp
;         DE              (SCRATCH REGISTER PAIR)
.sp
;         HL              LOAD ADDRESS
.sp
;         SP              SET TO STACK ADDRESS
.sp
;
START:    LXI  SP,3380H   ;SET STACK POINTER TO SCRATCH
                          ;AREA
          LXI  H,3380H    ;SET BASE LOAD ADDRESS
          MVI  B,0        ;START WITH TRACK 0
RDTRK:                    ;READ NEXT TRACK (INITIALLY 0)
          MVI  C,1        ;READ STARTING WITH SECTOR 1
.sp
RDSEC:                    ;READ NEXT SECTOR
          CALL READSEC    ;USER-SUPPLIED SUBROUTINE
          LXI  D,128      ;MOVE LOAD ADDRESS TO NEXT 1/2
                          ;PAGE
          DAD  D          ;HL = HL + 128
          INR  C          ;SECTOR = SECTOR + 1
          MOV  A,C        ;CHECK FOR END OF TRACK
          CPI  27
          JC   RDSEC      ;CARRY GENERATED IF SECTOR <27
.sp
;
;  ARRIVE HERE AT END OF TRACK, MOVE TO NEXT TRACK
          INR  B
          MOV  A,B        ;TEST FOR LAST TRACK
          CPI  2
          JC   RDTRK      ;CARRY GENERATED IF TRACK <2
.sp
;
;  USER-SUPPLIED SUBROUTINE TO READ THE DISK
READSEC:
;  ENTER WITH TRACK NUMBER IN REGISTER B,
          SECTOR NUMBER IN REGISTER C, AND
.sp
;         ADDRESS TO FILL IN HL
.sp
;
          PUSH B          ;SAVE B AND C REGISTERS
          PUSH H          ;SAVE HL REGISTERS
.sp 2
.sh
                  Listing 6-1.  GETSYS Program
.bp
          .................................................
          perform disk read at this point, branch to
          label START if an error occurs
          .................................................
          POP H           ;RECOVER HL
          POP B           ;RECOVER B AND C REGISTERS
          RET             ;BACK TO MAIN PROGRAM
.sp
          END START
.fi
.in 0
.sp 2
.sh
                    Listing 6-1.  (continued)
.sp 2
.pp
This program is assembled and listed in Appendix B for reference 
purposes, with an assumed origin of 100H.  The hexadecimal 
operation codes that are listed on the left might be useful if the 
program has to be entered through the panel switches.
.pp
The PUTSYS program can be constructed from GETSYS by changing 
only a few operations in the GETSYS program given above, as shown 
in Appendix C.  The register pair HL becomes the dump address, 
next address to write, and operations on these registers do not 
change within the program.  The READSEC subroutine is replaced by 
a WRITESEC subroutine, which performs the opposite function; data
from address HL is written to the track given by register B 
and sector given by register C.  It is often useful to combine 
GETSYS and PUTSYS into a single program during the test and 
development phase, as shown in Appendix C.
.sp 2
.tc    6.5  Disk Organization
.he CP/M Operating System Manual               6.5  Disk Organization
.sh
6.5  Disk Organization
.qs
.pp
The sector allocation for the standard distribution version of 
CP/M is given here for reference purposes.  The first sector contains
an optional software boot section (see the table on the following 
page.   Disk controllers are often set up to bring track 0, 
sector 1, into memory at a specific location, often location 
0000H.  The program in this sector, called BOOT, has the 
responsibility of bringing the remaining sectors into memory 
starting at location 3400H+b.  If the controller does not 
have a built-in sector load, the program in track 0, sector 1 can 
be ignored.  In this case, load the program from track 0, sector 
2, to location 3400H+b.
.pp
As an example, the Intel Model 800 
hardware cold start loader brings track 0, sector 1, into 
absolute address 3000H.  Upon loading this sector, control 
transfers to location 3000H, where the bootstrap operation 
commences by loading the remainder of track 0 and all of track 1 
into memory, starting at 3400H+b.  Note that this bootstrap 
loader is of little use in a non-microcomputer development system 
environment, although it is useful to examine it because some of 
the boot actions will have to be duplicated in the user's cold 
start loader.
.bp
.sh
             Table 6-3.  CP/M Disk Sector Allocation
.sp
.nf
Track #    Sector   Page#   Memory Address    CP/M Module name
.sp
  00         01             (boot address)    Cold Start Loader
  00         02      00        3400H+b               CCP
  '          03      '         3480H+b                '
  '          04      01        3500H+b                '
  '          05      '         3580H+b                '
  '          06      02        3600H+b                '
  '          07      '         3680H+b                '
  '          08      03        3700H+b                '
  '          09      '         3780H+b                '
  '          10      04        3800H+b                '
  '          11      '         3880H+b                '
  '          12      05        3900H+b                '
  '          13      '         3980H+b                '
  '          14      06        3A00H+b                '
  '          15      '         3A80H+b                '
  '          16      07        3B00H+b                '
  00         17      '         3B80H+b               CCP
  00         18      08        3C00H+b              BDOS
  '          19      '         3C80H+b                '
  '          20      09        3D00H+b                '
  '          21      '         3D80H+b                '
  '          22      10        3E00H+b                '
  '          23      '         3E80H+b                '
  '          24      11        3F00H+b                '
  '          25      '         3F80H+b                '
  '          26      12        4000H+b                '
  01         01      '         4080H+b                '
  '          02      13        4100H+b                '
  '          03      '         4180H+B                '
  '          04      14        4200H+b                '
  '          05      '         4280H+b                '
  '          06      15        4300H+b                '
  '          07      '         4380H+b                '
  '          08      16        4400H+b                '
  '          09      '         4480H+b                '
  '          10      17        4500H+b                '
  '          11      '         4580H+b                '
  '          12      18        4600H+b                '
  '          13      '         4680H+b                '
  '          14      19        4700H+b                '
  '          15      '         4780H+b                '
  '          16      20        4800H+b                '
  '          17      '         4880H+b                '
  '          18      21        4900H+b                '
.mb 4
.fm 1
  01         19      '         4900H+b              BDOS
  07         20      22        4A00H+b              BIOS
  '          21      '         4A80H+b                '
  '          22      23        4B00H+b                '
  '          23      '         4B80H+b                '
  '          24      24        4C00H+b                '
  01         25      '         4C80H+b              BIOS
  01         26      25        4D00H+b              BIOS
02-76      01-26                            (directory and data)
.fi
.nx sixb

















