.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                I-%
.pc 1
.tc I  CP/M Error Messages
.ce 2
.sh
Appendix I
.sp
.sh
CP/M Error Messages
.qs
.he CP/M Operating System Manual               I  CP/M Error Messages
.sp 2
.pp
Messages come from several different sources.  CP/M displays 
error messages when there are errors in calls to the Basic Disk 
Operating System (BDOS).  CP/M also displays messages when there 
are errors in command lines.  Each utility supplied with CP/M has 
its own set of messages.  The following lists CP/M messages and 
utility messages.  One might see messages other than those listed 
here if one is running an application program.  Check the 
application program's documentation for explanations of those 
messages.
.sp 2
.sh
                 Table I-1.  CP/M Error Messages
.sp
.ll 60
.nf
     Message        Meaning
.sp
.fi
.in 20
.ti -15
?
.sp
DDT.  This message has four possible meanings:
.sp
.in 23
.ti -2
o DDT does not understand the assembly language instruction.
.ti -2
o The file cannot be opened.
.ti -2
o A checksum error occurred in a HEX file.
.ti -2
o The assembler/disassembler was overlayed.
.sp 2
.in 20
.ti -15
ABORTED
.sp
PIP.  You stopped a PIP operation by pressing a key.
.sp 2
.ti -15
ASM Error Messages
.sp
.in 24
.ti -4
D   Data error:  data statement element cannot be placed in 
specified data area.
.sp
.ti -4
E   Expression error:  expression cannot be evaluated during 
assembly.
.sp
.ti -4
L   Label error:  label cannot appear in this context (might be 
duplicate label).
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.sp
     ASM Error Messages (continued)
.fi
.sp
.in 24
.ti -4
N   Not implemented:  unimplemented features, such as macros, are 
trapped.
.sp
.ti -4
O   Overflow:  expression is too complex to evaluate.
.sp
.ti -4
P   Phase error:  label value changes on two passes through 
assembly.
.sp
.ti -4
R   Register error:  the value specified as a register is 
incompatible with the code.
.sp
.ti -4
S   Syntax error:  improperly formed expression.
.sp
.ti -4
U   Undefined label:  label used does not exist.
.sp
.ti -4
V   Value error:  improperly formed operand encountered in an 
expression.
.sp 2
.in 20
.ti -15
BAD DELIMITER
.sp
STAT.  Check command line for typing errors.
.sp 2
.ti -15
Bad Load
.sp
CCP error message, or SAVE error message.
.sp 2
.ti -15
Bdos Err On d:
.sp
Basic Disk Operating System error on the designated drive:  CP/M 
replaces d: with the drive specification of the drive where the 
error occurred.  This message is followed by one of the four 
phrases in the situations described below.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
Bdos Err On d: Bad Sector
.sp
This message appears when CP/M finds no disk in the drive, when 
the disk is improperly formatted, when the drive latch is open, 
or when power to the drive is off.  Check for one of these 
situations and try again.  This could also indicate a hardware 
problem or a worn or improperly formatted disk.  Press ^C to 
terminate the program and return to CP/M, or press RETURN 
to ignore the error.
.sp 2
.ti -15
Bdos Err On d: File R/O
.sp
You tried to erase, rename, or set file attributes on a Read-Only 
file.  The file should first be set to Read-Write (R/W) with the 
command:  STAT filespec $R/W.
.sp 2
.ti -15
Bdos Err On d: R/O
.sp
Drive has been assigned Read-Only status with a STAT command, or 
the disk in the drive has been changed without being initialized 
with a ^C.  CP/M terminates the current program as soon as you 
press any key.
.sp 2
.ti -15
Bdos Err on d: Select
.sp
CP/M received a command line specifying a nonexistent drive.  
CP/M terminates the current program as soon as you press any key.  
Press RETURN or CTRL-C to recover.
.sp 2
.ti -15
Break "x" at c
.sp
ED.  "x" is one of the symbols described below and c is the 
command letter being executed when the error occurred.
.sp
.in 24
.ti -4
#   Search failure.  ED cannot find the string specified in an F, 
S, or N command.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 24
.ti -4
?   Unrecognized command letter c.  ED does not recognize the 
indicated command letter, or an E, H, Q, or O command is not 
alone on its command line.
.sp
.ti -4
O   The file specified in an R command cannot be found.
.sp
.ti -4
>   Buffer full.  ED cannot put any more characters in the memory 
buffer, or the string specified in an F, N, or S command is too 
long.
.sp
.ti -4
E   Command aborted.  A keystroke at the console aborted command 
execution.
.sp
Break "x" at c (continued)
.sp
.ti -4
F   Disk or directory full.  This error is followed by either the 
disk or directory full message.  Refer to the recovery procedures 
listed under these messages.
.sp 2
.in 20
.ti -15
CANNOT CLOSE DESTINATION FILE--{filespec}
.sp
PIP.  An output file cannot be closed.  You should take 
appropriate action after checking to see if the correct disk is 
in the drive and that the disk is not write-protected.
.sp 2
.nf
.in 5
Cannot close, R/O
CANNOT CLOSE FILES
.fi
.in 20
.sp
CP/M cannot write to the file.  This usually occurs because the 
disk is write-protected.
.sp
ASM.  An output file cannot be closed.  This is a fatal error 
that terminates ASM execution.  Check to see that the disk is in 
the drive, and that the disk is not write-protected.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
DDT.  The disk file written by a W command cannot be closed.  
This is a fatal error that terminates DDT execution.  Check if 
the correct disk is in the drive and that the disk is not write-protected.
.sp
SUBMIT.  This error can occur during SUBMIT file processing.  
Check if the correct system disk is in the A drive and that the 
disk is not write-protected.  The SUBMIT job can be restarted 
after rebooting CP/M.
.sp 2
.ti -15
CANNOT READ
.sp
PIP.  PIP cannot read the specified source.  Reader cannot be 
implemented.
.sp 2
.ti -15
CANNOT WRITE
.sp
PIP.  The destination specified in the PIP command is illegal.  
You probably specified an input device as a destination.
.sp 2
.ti -15
Checksum error
.sp
PIP.  A HEX record checksum error was encountered.  The HEX 
record that produced the error must be corrected, probably by 
recreating the HEX file.
.sp 2
.nf
.in 5
CHECKSUM ERROR
LOAD ADDRESS hhhh
ERROR ADDRESS hhhh
BYTES READ:
hhhh:
.fi
.in 20
.sp
LOAD.  File contains incorrect data.  Regenerate HEX file from 
the source.
.sp 2
.ti -15
Command Buffer Overflow
.sp
SUBMIT.  The SUBMIT buffer allows up to 2048 characters in the 
input file.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
Command too long
.sp
SUBMIT.  A command in the SUBMIT file cannot exceed 125 
characters.
.sp 2
.ti -15
CORRECT ERROR, TYPE RETURN OR CTRL-Z
.sp
PIP.  A HEX record checksum was encountered during the transfer 
of a HEX file.  The HEX file with the checksum error should be 
corrected, probably by recreating the HEX file.
.sp 2
.ti -15
DESTINATION IS R/O, DELETE (Y/N)?
.sp
PIP.  The destination file specified in a PIP command already 
exists and it is Read-Only.  If you type Y, the destination file 
is deleted before the file copy is done.
.sp 2
.ti -15
Directory full
.sp
ED.  There is not enough directory space for the file being 
written to the destination disk.  You can use the OXfilespec 
command to erase any unnecessary files on the disk without 
leaving the editor.
.sp
SUBMIT.  There is not enough directory space to write the $$$.SUB 
file used for processing SUBMITs.  Erase some files or select a 
new disk and retry.
.sp 2
.ti -15
Disk full
.sp
ED.  There is not enough disk space for the output file.  This 
error can occur on the W, E, H, or X commands.  If it occurs with 
X command, you can repeat the command prefixing the filename with 
a different drive.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
DISK READ ERROR--{filespec}
.sp
PIP.  The input disk file specified in a PIP command cannot be 
read properly.  This is usually the result of an unexpected end-of-file.
Correct the problem in your file.
.sp 2
.ti -15
DISK WRITE ERROR--{filespec}
.sp
DDT.  A disk write operation cannot be successfully performed 
during a W command, probably due to a full disk.  You should 
either erase some unnecessary files or get another disk with more 
space.
.sp
PIP.  A disk write operation cannot be successfully performed 
during a PIP command, probably due to a full disk.  You should 
either erase some unnecessary files or get another disk with more 
space and execute PIP again.
.sp
SUBMIT.  The SUBMIT program cannot write the $$$.SUB file to the 
disk.  Erase some files, or select a new disk and try again.
.sp 2
.ti -15
ERROR: BAD PARAMETER
.sp
PIP.  You entered an illegal parameter in a PIP command.  Retype 
the entry correctly.
.sp 2
.ti -15
ERROR: CANNOT OPEN SOURCE, LOAD ADDRESS hhhh
.sp
LOAD.  Displayed if LOAD cannot find the specified file or if no 
filename is specified.
.sp 2
.ti -15
ERROR: CANNOT CLOSE FILE, LOAD ADDRESS hhhh
.sp
LOAD.  Caused by an error code returned by a BDOS function call.  
Disk might be write-protected.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
ERROR: CANNOT OPEN SOURCE, LOAD ADDRESS hhhh
.sp
LOAD.  Cannot find source file.  Check disk directory.
.sp 2
.ti -15
ERROR: DISK READ, LOAD ADDRESS hhhh
.sp
LOAD.  Caused by an error code returned by a BDOS function call.
.sp 2
.ti -15
ERROR: DISK WRITE, LOAD ADDRESS hhhh
.sp
LOAD.  Destination disk is full.
.sp 2
.ti -15
ERROR: INVERTED LOAD ADDRESS, LOAD ADDRESS hhhh
.sp
LOAD.  The address of a record was too far from the address of 
the previously-processed record.  This is an internal limitation 
of LOAD, but it can be circumvented.  Use DDT to read the HEX 
file into memory, then use a SAVE command to store the memory 
image file on disk.
.sp 2
.ti -15
ERROR: NO MORE DIRECTORY SPACE, LOAD ADDRESS hhhh
.sp
LOAD.  Disk directory is full.
.sp 2
.ti -15
Error on line nnn message
.sp
SUBMIT.  The SUBMIT program displays its messages in the format 
shown above, where nnn represents the line number of the SUBMIT 
file.  Refer to the message following the line number.
.sp 2
.ti -15
FILE ERROR
.sp
ED.  Disk or directory is full, and ED cannot write anything more 
on the disk.  This is a fatal error, so make sure there is enough 
space on the disk to hold a second copy of the file before 
invoking ED.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
FILE EXISTS
.sp
You have asked CP/M to create or rename a file using a file 
specification that is already assigned to another file.  Either 
delete the existing file or use another file specification.
.sp
REN.  The new name specified is the name of a file that already 
exists.  You cannot rename a file with the name of an existing 
file.  If you want to replace an existing file with a newer 
version of the same file, either rename or erase the existing 
file, or use the PIP utility.
.sp 2
.ti -15
File exists, erase it
.sp
ED.  The destination filename already exists when you are placing 
the destination file on a different disk than the source.  It 
should be erased or another disk selected to receive the output 
file.
.sp 2
.ti -15
** FILE IS READ/ONLY **
.sp
ED.  The file specified in the command to invoke ED has the
Read-Only attribute.  Ed can read the file so that the user can 
examine it, but ED cannot change a Read-Only file.
.sp 2
.mb 4
.fm 1
.ti -15
File Not Found
.sp
CP/M cannot find the specified file.  Check that you have entered 
the correct drive specification or that you have the correct disk 
in the drive.
.sp
ED.  ED cannot find the specified file.  Check that you have 
entered the correct drive specification or that you have the 
correct disk in the drive.
.sp
STAT.  STAT cannot find the specified file.  The message might 
appear if you omit the drive specification.  Check to see if the 
correct disk is in the drive.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
FILE NOT FOUND--{filespec}
.sp
.mb 6
.fm 2
PIP.  An input file that you have specified does not exist.
.sp 2
.ti -15
Filename required
.sp
ED.  You typed the ED command without a filename.  Reenter the ED 
command followed by the name of the file you want to edit or 
create.
.sp 2
.ti -15
hhhh??=dd
.sp
DDT.  The ?? indicates DDT does not know how to represent the 
hexadecimal value dd encountered at address hhhh in 8080 assembly 
language.  dd is not an 8080 machine instruction opcode.
.sp 2
.ti -15
Insufficient memory
.sp
DDT.  There is not enough memory to load the file specified in an 
R or E command.
.sp 2
.ti -15
Invalid Assignment
.sp
STAT.  You specified an invalid drive or file assignment, or 
misspelled a device name.  This error message might be followed 
by a list of the valid file assignments that can follow a 
filename.  If an invalid drive assignment was attempted the 
message Use: d:=RO is displayed, showing the proper syntax for 
drive assignments.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
Invalid control character
.sp
SUBMIT.  The only valid control characters in the SUBMIT files of 
the type SUB are ^ A through ^ Z.  Note that in a SUBMIT file the 
control character is represented by typing the circumflex, ^, not 
by pressing the control key.
.sp 2
.ti -15
INVALID DIGIT--{filespec}
.sp
PIP.  An invalid HEX digit has been encountered while reading a 
HEX file.  The HEX file with the invalid HEX digit should be 
corrected, probably by recreating the HEX file.
.sp 2
.ti -15
Invalid Disk Assignment
.sp
STAT.  Might appear if you follow the drive specification with 
anything except =R/O.
.sp 2
.ti -15
INVALID DISK SELECT
.sp
CP/M received a command line specifying a nonexistent drive, or 
the disk in the drive is improperly formatted.  CP/M terminates 
the current program as soon as you press any key.
.sp 2
.ti -15
INVALID DRIVE NAME (Use A, B, C, or D)
.sp
SYSGEN.  SYSGEN recognizes only drives A, B, C, and D as valid 
destinations for system generation.
.sp 2
.ti -15
Invalid File Indicator
.sp
STAT.  Appears if you do not specify RO, RW, DIR, or SYS.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
INVALID FORMAT
.sp
PIP.  The format of your PIP command is illegal.  See the 
description of the PIP command.
.sp 2
.nf
.in 5
INVALID HEX DIGIT
LOAD ADDRESS hhhh
ERROR ADDRESS hhhh
BYTES READ:
hhhh
.fi
.in 20
.sp
LOAD.  File contains incorrect HEX digit.
.sp 2
.ti -15
INVALID MEMORY SIZE
.sp
MOVCPM.  Specify a value less than 64K or your computer's actual 
memory size.
.sp 2
.ti -15
INVALID SEPARATOR
.sp
PIP.  You have placed an invalid character for a separator 
between two input filenames.
.sp 2
.ti -15
INVALID USER NUMBER
.sp
PIP.  You have specified a user number greater than 15.  User 
numbers are in the range 0 to 15.
.sp 2
.ti -15
n?
.sp
USER.  You specified a number greater than fifteen for a user 
area number.  For example, if you type USER 18<cr>, the screen 
displays 18?.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
NO DIRECTORY SPACE
.sp
ASM.  The disk directory is full.  Erase some files to make room 
for PRN and HEX files.  The directory can usually hold only 64 
filenames.
.sp 2
.ti -15
NO DIRECTORY SPACE--{filespec}
.sp
PIP.  There is not enough directory space for the output file.  
You should either erase some unnecessary files or get another 
disk with more directory space and execute PIP again.
.sp 2
.ti -15
NO FILE--{filespec}
.sp
DIR, ERA, REN, PIP.  CP/M cannot find the specified file, or no 
files exist.
.sp
ASM.  The indicated source or include file cannot be found on the 
indicated drive.
.sp
DDT.  The file specified in an R or E command cannot be found on 
the disk.
.sp 2
.ti -15
NO INPUT FILE PRESENT ON DISK
.sp
DUMP.  The file you requested does not exist.
.sp 2
.ti -15
No memory
.sp
There is not enough (buffer?) memory available for loading the 
program specified.
.sp 2
.ti -15
NO SOURCE FILE ON DISK
.sp
SYSGEN.  SYSGEN cannot find CP/M either in CPMxx.com form or on 
the system tracks of the source disk.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
NO SOURCE FILE PRESENT
.sp
ASM.  The assembler cannot find the file you specified.  Either 
you mistyped the file specification in your command line, or the 
filetype is not ASM.
.sp 2
.ti -15
NO SPACE
.sp
SAVE.  Too many files are already on the disk, or no room is left 
on the disk to save the information.
.sp 2
.ti -15
No SUB file present
.sp
SUBMIT.  For SUBMIT to operate properly, you must create a file 
with filetype of SUB.  The SUB file contains usual CP/M commands.  
Use one command per line.
.sp 2
.ti -15
NOT A CHARACTER SOURCE
.sp
PIP.  The source specified in your PIP command is illegal.  You 
have probably specified an output device as a source.
.sp 2
.ti -15
** NOT DELETED **
.sp
PIP.  PIP did not delete the file, which might have had the R/O 
attribute.
.sp 2
.ti -15
NOT FOUND
.sp
PIP.  PIP cannot find the specified file.
.sp 2
.ti -15
OUTPUT FILE WRITE ERROR
.sp
ASM.  You specified a write-protected disk as the destination for 
the PRN and HEX files, or the disk has no space left.  Correct 
the problem before assembling your program.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
Parameter error
.sp
SUBMIT.  Within the SUBMIT file of type sub, valid parameters are 
$0 through $9.
.sp 2
.ti -15
PARAMETER ERROR, TYPE RETURN TO IGNORE
.sp
SYSGEN.  If you press RETURN, SYSGEN proceeds without processing 
the invalid parameter.
.sp 2
.ti -15
QUIT NOT FOUND
.sp
PIP.  The string argument to a Q parameter was not found in your 
input file.
.sp 2
.ti -15
Read error
.sp
TYPE.  An error occurred when reading the file specified in the 
type command.  Check the disk and try again.  The STAT filespec 
command can diagnose trouble.
.sp 2
.ti -15
READER STOPPING
.sp
PIP.  Reader operation interrupted.
.sp 2
.ti -15
Record Too Long
.sp
PIP.  PIP cannot process a record longer than 128 bytes.
.sp 2
.ti -15
Requires CP/M 2.0 or later
.sp
XSUB.  XSUB requires the facilities of CP/M 2.0 or newer version.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
Requires CP/M 2.0 or new for operation
.sp
PIP.  This version of PIP requires the facilities of CP/M 2.0 or 
newer version.
.sp 2
.ti -15
START NOT FOUND
.sp
PIP.  The string argument to an S parameter cannot be found in 
the source file.
.sp 2
.ti -15
SOURCE FILE INCOMPLETE
.sp
SYSGEN.  SYSGEN cannot use your CP/M source file.
.sp 2
.ti -15
SOURCE FILE NAME ERROR
.sp
ASM.  When you assemble a file, you cannot use the wildcard 
characters * and ? in the filename. Only one file can be 
assembled at a time.
.sp 2
.ti -15
SOURCE FILE READ ERROR
.sp
ASM.  The assembler cannot understand the information in the file 
containing the assembly-language program.  Portions of another 
file might have been written over your assembly-language file, or 
information was not properly saved on the disk.  Use the TYPE 
command to locate the error.  Assembly-language files contain the 
letters, symbols, and numbers that appear on your keyboard.  If 
your screen displays unrecognizable output or behaves strangely, 
you have found where computer instructions have crept into your 
file.
.sp 2
.ti -15
SYNCHRONIZATION ERROR
.sp
MOVCPM.  The MOVCPM utility is being used with the wrong CP/M 
system.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
"SYSTEM" FILE NOT ACCESSIBLE
.sp
You tried to access a file set to SYS with the STAT command.
.sp 2
.ti -15
** TOO MANY FILES **
.sp
STAT.  There is not enough memory for STAT to sort the files 
specified, or more than 512 files were specified.
.sp 2
.ti -15
UNEXPECTED END OF HEX FILE--{filespec}
.sp
PIP.  An end-of-file was encountered prior to a termination HEX 
record.  The HEX file without a termination record should be 
corrected, probably by recreating the HEX file.
.sp 2
.ti -15
Unrecognized Destination
.sp
PIP.  Check command line for valid destination.
.sp 2
.ti -15
Use: STAT d:=RO
.sp
STAT.  An invalid STAT drive command was given.  The only valid 
drive assignment in STAT is STAT d:=RO.
.sp 2
.ti -15
VERIFY ERROR:--{filespec}
.sp
PIP.  When copying with the V option, PIP found a difference when 
rereading the data just written and comparing it to the data in 
its memory buffer.  Usually this indicates a failure of either 
the destination disk or drive.
.sp 2
.ti -15
WRONG CP/M VERSION (REQUIRES 2.0)
.sp 2
.ti -15
XSUB ACTIVE
.sp
SUBMIT.  XSUB has been invoked.
.in 0
.bp
.sh
                     Table I-1.  (continued)
.sp
.nf
     Message        Meaning
.fi
.sp
.in 20
.ti -15
XSUB ALREADY PRESENT
.sp
SUBMIT.  XSUB is already active in memory.
.sp
.ti -15
Your input?
.sp
If CP/M cannot find the command you specified, it returns the 
command name you entered followed by a question mark.  Check that 
you have typed the command line correctly, or that the command 
you requested exists as a .COM file on the default or specified 
disk.
.in 0
.ll 65
.sp 2
.ce
End of Appendix I
