.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                1-%
.pc 1
.tc 1  CP/M Features and Facilities
.ce 2
.sh
Section 1
.sp
.sh
CP/M Features and Facilities
.qs
.sp 3
.he CP/M Operating System Manual                    1.1  Introduction
.tc    1.1  Introduction
.sh
1.1  Introduction
.qs
.fi
.pp 5
CP/M \ is a monitor control program for microcomputer system development that
uses floppy disks or Winchester hard disks for backup storage.  Using a
computer system based on the Intel \ 8080 microcomputer, CP/M provides an
environment for program construction, storage, and editing, along
with assembly and program check-out facilities.  CP/M can be easily
altered to execute with any computer
configuration that uses a Zilog \ Z80 \ or an Intel 8080 Central Processing
Unit (CPU) and has at least 20K bytes of main memory with up to 16 disk
drives.  A detailed discussion of the modifications required for any
particular hardware environment is given in Section 6.  Although the
standard Digital Research version operates on a single-density 
Intel Model 800, microcomputer development system several different 
hardware manufacturers support their own input-output (I/O) drivers for CP/M.
.pp
The CP/M monitor provides rapid access to programs through a comprehensive
file management package.  The file subsystem supports a named file structure,
allowing dynamic allocation of file space as well as sequential and random
file access.  Using this file system, a large number of programs can be
stored in both source and machine-executable form.
.pp
CP/M 2 is a high-performance, single console operating system that uses
table-driven techniques to allow field reconfiguration to match a wide
variety of disk capacities.  All fundamental file restrictions are removed,
maintaining upward compatibility from previous versions of release 1.
.pp
Features of CP/M 2 include field specification of one to sixteen logical
drives, each containing up to eight megabytes.  Any particular file can
reach the full drive size with the capability of expanding to thirty-two
megabytes in future releases.  The directory size can be field-configured to
contain any reasonable number of entries, and each file is optionally tagged
with Read-Only and system attributes.  Users of CP/M 2 are physically
separated by user numbers, with facilities for file copy operations from one
user area to another.  Powerful relative-record random access functions are
present in CP/M 2 that provide direct access to any of the 65536 records of
an eight-megabyte file.
.pp
CP/M also supports ED, a powerful context editor, ASM , an Intel-compatible
assembler, and DDT , debugger subsystems.  Optional software includes a
powerful
Intel-compatible macro assembler, symbolic debugger, along with various
high-level languages.  When coupled with CP/M's Console Command 
Processor (CCP),
the resulting facilities equal or exceed similar large computer facilities.
.pp
CP/M is logically divided into several distinct parts:
.sp
.in 3
.nf
o BIOS (Basic I/O System), hardware-dependent
o BDOS (Basic Disk Operating System)
o CCP (Console Command Processor)
o TPA (Transient Program Area)
.fi
.in 0
.pp
The BIOS provides the primitive operations necessary to access the disk
drives and to interface standard peripherals:  teletype, CRT, paper tape
reader/punch, and user-defined peripherals.  You can tailor 
peripherals for any particular hardware environment by patching this
portion of
CP/M.  The BDOS provides disk management by controlling one or more disk
drives containing independent file directories.  The BDOS implements disk
allocation strategies that provide fully dynamic file construction while
minimizing head movement across the disk during access.  The BDOS has entry
points that include the following primitive operations, which the
program accesses:
.sp
.in 5
.ti -2
o SEARCH looks for a particular disk file by name.
.ti -2
o OPEN opens a file for further operations.
.ti -2
o CLOSE closes a file after processing.
.ti -2
o RENAME changes the name of a particular file.
.ti -2
o READ reads a record from a particular file.
.ti -2
o WRITE writes a record to a particular file.
.ti -2
o SELECT selects a particular disk drive for further operations.
.in 0
.pp
The CCP provides a symbolic interface between your console and the
remainder of the CP/M system.  The CCP reads the console device and
processes commands, which include listing the file directory, printing the
contents of files, and controlling the operation of transient programs, such
as assemblers, editors, and debuggers.  The standard commands that are
available in the CCP are listed in Section 1.2.1.
.pp
The last segment of CP/M is the area called the Transient Program
Area (TPA).  The TPA holds programs that are loaded from the disk under
command of the CCP.  During program editing, for example, the TPA holds
the CP/M text editor machine code and data areas.  Similarly, programs
created under CP/M can be checked out by loading and executing these
programs in the TPA.
.pp
Any or all of the CP/M component subsystems can be overlaid by an
executing program.  That is, once a user's program is loaded into the TPA,
the CCP, BDOS, and BIOS areas can be used as the program's data area.
A bootstrap loader is programmatically accessible whenever the BIOS portion
is not overlaid; thus, the user program need only branch to the bootstrap
loader at the end of execution and the complete CP/M monitor is reloaded
from disk.
.pp
The CP/M operating system is partitioned into distinct modules, including
the BIOS portion that defines the hardware environment in which CP/M is
executing.  Thus, the standard system is easily modified to any nonstandard
environment by changing the peripheral drivers to handle the custom system.
.bp
.tc    1.2  Functional Description
.he CP/M Operating System Manual          1.2  Functional Description
.sh
1.2  Functional Description
.qs
.pp
You interact with CP/M primarily through the CCP, which reads and
interprets commands entered through the console.  In general, the CCP
addresses one of several disks that are on-line.  The standard system
addresses up to sixteen different disk drives.  These disk drives are
labeled A through P.  A disk is logged-in if the CCP is currently
addressing the disk.  To clearly indicate which disk is the currently logged
disk, the CCP always prompts the operator with the disk name followed by the
symbol >, indicating that the CCP is ready for another command.  Upon
initial start-up, the CP/M system is loaded from disk A, and the CCP
displays the following message:
.sp
.ti 8
CP/M VER x.x
.sp
where x.x is the CP/M version number.  All CP/M systems are initially set
to operate in a 20K memory space, but can be easily reconfigured to fit any
memory size on the host system (see Section 1.6.9).  Following system
sign-on, CP/M automatically logs in disk A, prompts you with the
symbol A>, indicating that CP/M is currently addressing disk A, and
waits for a command.  The commands are implemented at two levels:  built-in
commands and transient commands.
.sp 2
.tc         1.2.1  General Command Structure
.sh
1.2.1  General Command Structure
.qs
.pp
Built-in commands are a part of the CCP program, while transient
commands are loaded into the TPA from disk and executed.  The 
following are built-in commands:
.sp
.in 3
.nf
o ERA erases specified files.
o DIR lists filenames in the directory.
o REN renames the specified file.
o SAVE saves memory contents in a file.
o TYPE types the contents of a file on the logged disk.
.in 0
.fi
.sp
Most of the commands reference a particular file or group of files.  The
form of a file reference is specified in Section 1.2.2.
.sp 2
.tc         1.2.2  File References
.sh
1.2.2  File References
.qs
.pp
A file reference identifies a particular file or group of files on a
particular disk attached to CP/M.  These file references are
either unambiguous (ufn) or ambiguous (afn).  An unambiguous file
reference uniquely identifies a single file, while an ambiguous file
reference is satisfied by a number of different files.
.mb 5
.fm 1
.pp
File references consist of two parts:  the primary filename and the
filetype.  Although the filetype is optional, it usually is 
generic.  For example, the filetype ASM is used to denote that the file is an
assembly language source file, while the primary filename distinguishes each
particular source file.  The two names are separated by a period, as shown
in the following example:
.bp
.ti 8
filename.typ
.sp
.mb 6
.fm 2
In this example, filename is the primary filename of eight characters or
less, and typ
is the filetype of no more than three characters.  As mentioned above, the
name
.sp
.ti 8
filename
.sp
is also allowed and is equivalent to a filetype consisting of 
three blanks.  The characters used in specifying an unambiguous 
file reference cannot contain any of the following special 
characters:
.sp
.ti 8
< > . , ; : = ? * [ ] _ % | ( ) / \\
.sp
while all alphanumerics and remaining special characters are allowed.
.pp
An ambiguous file reference is used for directory search and pattern
matching.  The form of an ambiguous file reference is similar to an
unambiguous reference, except the symbol ? can be interspersed throughout
the primary and secondary names.  In various commands throughout CP/M,
the ? symbol matches any character of a filename in the ? position.
Thus, the ambiguous reference
.sp
.ti 8
X?Z.C?M
.sp
matches the following unambiguous filenames
.sp
.ti 8
XYZ.COM
.sp
and
.sp
.ti 8
X3Z.CAM
.sp
The * wildcard character can also be used in an ambiguous file 
reference.  The * character replaces all or part of a filename or 
filetype.  Note that
.sp
.ti 8
*.*
.sp
equals the ambiguous file reference
.sp
.ti 8
????????.???
.sp
while
.sp
.ti 8
filename.*
.sp
and
.sp
.ti 8
*.typ
.sp
are abbreviations for
.sp
.ti 8
filename.???
.sp
and
.sp
.ti 8
????????.typ
.sp
respectively.  As an example,
.sp
.ti 8
A>\c
.sh
DIR *.*
.qs
.sp
is interpreted by the CCP as a command to list the names of all disk files
in the directory.  The following example searches only for a file 
by the name X.Y:
.sp
.ti 8
A>\c
.sh
DIR X,Y
.qs
.sp
Similarly, the command
.sp
.ti 8
A>\c
.sh
DIR X?Y.C?M
.qs
.sp
causes a search for all unambiguous filenames on the disk that satisfy
this ambiguous reference.
.pp
The following file references are valid unambiguous file references:
.sp
.nf
.in 8
X
X.Y
XYZ
XYZ.COM
GAMMA
GAMMA.1
.fi
.in 0
.pp
As an added convenience, the programmer can generally specify the disk drive
name along with the filename.  In this case, the drive name is given as a
letter A through P followed by a colon (:).  The specified drive is
then logged-in before the file operation occurs.  Thus, the following are
valid file references with disk name prefixes:
.sp
.nf
.in 8
A:X.Y
P:XYZ.COM
B:XYZ
B:X.A?M
C:GAMMA
C:*.ASM
.fi
.in 0
.sp
All alphabetic lower-case letters in file and drive names are translated to
upper-case when they are processed by the CCP.
.sp 2
.tc    1.3  Switching Disks
.he CP/M Operating System Manual                 1.3  Switching Disks
.sh
1.3  Switching Disks
.qs
.mb 5
.fm 1
.pp
The operator can switch the currently logged disk by typing the disk drive
name, A through P, followed by a colon when the CCP is waiting for
console input.  The following sequence of prompts and commands
can occur after the CP/M system is loaded from disk A:
.sp
.nf
.in 8
CP/M VER 2.2
A>\c
.sh
DIR                     \c
.qs
List all files on disk A.
A:SAMPLE ASM SAMPLE PRN
A>\c
.sh
B:                      \c
.qs
Switch to disk B.
B>\c
.sh
DIR *.ASM               \c
.qs
List all ASM files on B.
B:DUMP ASM FILES ASM
b>\c
.sh
A:                      \c
.qs
Switch back to A.
.in 0
.fi
.mb 6
.fm 2
.sp 2
.tc    1.4  Built-in Commands
.he CP/M Operating System Manual               1.4  Built-in Commands
.sh
1.4  Built-in Commands
.qs
.pp
The file and device reference forms described can now be used to fully
specify the structure of the built-in commands.  Assume the following
abbreviations in the description below:
.sp
.in 8
.nf
ufn   unambiguous file reference
afn   ambiguous file reference
.fi
.in 0
.sp
Recall that the CCP always translates lower-case characters to upper-case
characters internally.  Thus, lower-case alphabetics are treated as if they
are upper-case in command names and file references.
.sp 2
.tc         1.4.1  ERA Command
.sh
1.4.1  ERA Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
ERA afn
.pp
The ERA (erase) command removes files from the currently logged-in
disk, for example, the disk name currently prompted by CP/M preceding the >.
The files that are erased are those that satisfy the ambiguous file
reference afn.  The following examples illustrate the use of ERA:
.sp 2
.in 24
.ti -16
ERA X.Y         The file named X.Y on the currently logged disk is removed
from the disk directory and the space is returned.
.sp
.ti -16
ERA X.*         All files with primary name X are removed from the current
disk.
.sp
.ti -16
ERA *.ASM       All files with secondary name ASM are removed from the
current disk.
.sp
.ti -16
ERA X?Y.C?M     All files on the current disk that satisfy the ambiguous
reference X?Y.C?M are deleted.
.bp
.ti -16
ERA *.*         Erase all files on the current disk.  In this 
case, the CCP prompts the console with the message
.sp
.nf
ALL FILES (Y/N)?
.fi
.sp
which requires a Y response before files are actually removed.
.sp
.ti -16
ERA b:*.PRN     All files on drive B that satisfy the ambiguous
reference ????????.PRN are deleted, independently of the currently
logged disk.
.in 0
.sp 3
.tc         1.4.2  DIR Command
.sh
1.4.2  DIR Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
DIR afn
.pp
The DIR (directory) command causes the names of all files that satisfy the
ambiguous filename afn to be listed at the console device.  As a special
case, the command
.sp
.ti 8
DIR
.sp
lists the files on the currently logged disk (the command DIR is
equivalent to the command DIR *.*).  The following are valid DIR 
commands:
.sp
.nf
.in 8
DIR X.Y
DIR X?Z.C?M
DIR ??.Y
.in 0
.fi
.pp
Similar to other CCP commands, the afn can be preceded by a drive name.
The following DIR commands cause the selected drive to be addressed before
the directory search takes place:
.sp
.in 8
.nf
DIR B:
DIR B:X.Y
DIR B:*.A?M
.fi
.in 0
.pp
If no files on the selected disk satisfy the directory request, the
message NO FILE appears at the console.
.bp
.tc         1.4.3  REN Command
.sh
1.4.3  REN Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
REN ufn1=ufn2
.pp
The REN (rename) command allows you to change the names of files on
disk.  The file satisfying ufn2 is changed to ufn1.  The currently logged
disk is assumed to contain the file to rename (ufn2).  You can also
type a left-directed arrow instead of the equal sign if the console supports
this graphic character.  The following are examples of the REN 
command:
.sp 2
.in 31
.ti -23
REN X.Y=Q.R            The file Q.R is changed to X.Y.
.ti -23
.sp
REN XYZ.COM=XYZ.XXX    The file XYZ.XXX is changed to XYZ.COM.
.in 0
.fi
.sp
.pp
The operator precedes either ufn1 or ufn2 (or both) by an optional drive
address.  If ufn1 is preceded by a drive name, then ufn2 is assumed to exist
on the same drive.  Similarly, if ufn2 is preceded by a drive name, then
ufn1 is assumed to exist on the drive as well.  The same drive must be
specified in both cases if both ufn1 and ufn2 are preceded by drive names.
The following REN commands illustrate this format:
.sp 2
.in 31
.ti -23
REN A:X.ASM=Y.ASM      The file Y.ASM is changed to X.ASM on drive A.
.sp
.ti -23
REN B:ZAP.BAS=ZOT.BAS  The file ZOT.BAS is changed to ZAP.BAS on drive B.
.sp
.ti -23
REN B:A.ASM=B:A.BAK    The file A.BAK is renamed to A.ASM on drive B.
.in 0
.sp
.pp
If ufn1 is already present, the REN command responds with the
error FILE EXISTS and not perform the change.  If ufn2 does not exist on
the specified disk, the message NO FILE is printed at the console.
.sp 2
.tc         1.4.4  SAVE Command
.sh
1.4.4  SAVE Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
SAVE n ufn
.pp
The SAVE command places n pages (256-byte blocks) onto disk from the TPA
and names this file ufn.  In the CP/M distribution system, the TPA starts
at 100H (hexadecimal) which is the second page of memory.  The SAVE command
must specify 2 pages of memory if the user's program occupies the area
from 100H through 2FFH.  The machine code file can be subsequently loaded
and executed.  The following are examples of the SAVE command:
.sp 2
.in 31
.ti -23
SAVE 3X.COM            Copies 100H through 3FFH to X.COM.
.sp
.ti -23
SAVE 40 Q              Copies 100H through 28FFH to Q.  Note that 28 is the
page count in 28FFH, and that 28H = 2*16+8=40 decimal.
.sp
.ti -23
SAVE 4 X.Y             Copies 100H through 4FFH to X.Y.
.in 0
.sp 2
The SAVE command can also specify a disk drive in the ufn portion of the
command, as shown in the following example:
.sp
.in 31
.ti -23
SAVE 10 B:ZOT.COM      Copies 10 pages, 100H through 0AFFH, to the
file ZOT.COM on drive B.
.in 0
.sp 3
.tc         1.4.5  TYPE Command
.sh
1.4.5  TYPE Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
TYPE ufn
.pp
The TYPE command displays the content of the ASCII source file ufn on the
currently logged disk at the console device.  The following are valid
TYPE commands:
.sp
.in 8
.nf
TYPE X.Y
TYPE X.PLM
TYPE XXX
.in 0
.fi
.pp
The TYPE command expands tabs, CTRL-I characters, assuming tab positions are
set at every eighth column.  The ufn can also reference a drive name.
.sp
.in 24
.ti -16
TYPE B:X.PRN    The file X.PRN from drive B is displayed.
.in 0
.sp 2
.tc         1.4.6  USER Command
.sh
1.4.6  USER Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.ti 8
USER n
.pp
The USER command allows maintenance of separate files in the same 
directory.  In the syntax line, n is an integer value in the range 0 to
15.  On cold start, the operator is automatically logged into user
area number 0, which is
compatible with standard CP/M 1 directories.  You can issue the
USER command at any time to move to another logical area within the same
directory.  Drives that are logged-in while addressing one user number are
automatically active when the operator moves to another.  A user number is
simply a prefix that accesses particular directory entries on the active
disks.
.pp
The active user number is maintained until changed by a subsequent USER
command, or until a cold start when user 0 is again assumed.
.sp 2
.tc    1.5  Line Editing and Output Control
.he CP/M Operating System Manual 1.5  Line Editing and Output Control
.sh
1.5  Line Editing and Output Control
.pp
The CCP allows certain line-editing functions while typing command lines.
The CTRL-key sequences are obtained by pressing the control and letter keys
simultaneously.  Further, CCP command lines are generally up to 255
characters in length; they are not acted upon until the carriage return key
is pressed.
.sp 2
.ce
.sh
Table 1-1.  Line-editing Control Characters
.qs
.sp
.ll 60
.in 5
.nf
Character                      Meaning
.fi
.sp
.in 18
.ti -12
CTRL-C      Reboots CP/M system when pressed at start of line.
.sp
.ti -12
CTRL-E      Physical end of line; carriage is returned, but line is not sent
until the carriage return key is pressed.
.sp
.ti -12
CTRL-H      Backspaces one character position.
.sp
.ti -12
CTRL-J      Terminates current input (line feed).
.sp
.ti -12
CTRL-M      Terminates current input (carriage return).
.sp
.ti -12
CTRL-P      Copies all subsequent console output to the currently
assigned list device (see Section 1.6.1).  Output is sent to the list device
and the console device until the next CTRL-P is pressed.
.sp
.ti -12
CTRL-R      Retypes current command line; types a clean line following
character deletion with rubouts.
.sp
.ti -12
CTRL-S      Stops the console output temporarily.  Program execution and
output continue when you press any character at the console, for 
example another CTRL-S.  This feature stops output on high speed consoles,
such as CRTs, in order to view a segment of output before continuing.
.bp
.ll 65
.in 0
.ce
.sh
Table 1-1.  (continued)
.qs
.sp
.ll 60
.in 5
.nf
Character                      Meaning
.fi
.sp
.in 18
.ti-12
CTRL-U      Deletes the entire line typed at the console.
.sp
.ti -12
CTRL-X      Same as CTRL-U.
.sp
.ti -12
CTRL-Z      Ends input from the console (used in PIP and ED).
.sp
.ti -12
RUB/DEL     Deletes and echoes the last character typed at the console.
.in 0
.ll 65
.sp 2
.tc    1.6  Transient Commands
.he CP/M Operating System Manual              1.6  Transient Commands
.sh
1.6  Transient Commands
.qs
.pp
Transient commands are loaded from the currently logged disk and executed in
the TPA.  The transient commands for execution under the CCP are below.
Additional functions are easily defined by the user (see Section 1.6.3).
.sp 2
.ce
.sh
Table 1-2.  CP/M Transient Commands
.qs
.sp
.ll 60
.in 5
.nf
Command                      Function
.fi
.sp
.in 16
.ti -11
STAT       Lists the number of bytes of storage remaining on the currently
logged disk, provides statistical information about particular files, and
displays or alters device assignment.
.sp
.ti -11
ASM        Loads the CP/M assembler and assembles the specified program from
disk.
.sp
.ti -11
LOAD       Loads the file in Intel HEX machine code format and produces a
file in machine executable form which can be loaded into the TPA.  This loaded
program becomes a new command under the CCP.
.sp
.ti -11
DDT        Loads the CP/M debugger into TPA and starts execution.
.sp
.ti -11
PIP        Loads the Peripheral Interchange Program for subsequent disk file
and peripheral transfer operations.
.sp
.ti-11
ED         Loads and executes the CP/M text editor program.
.sp
.ti -11
SYSGEN     Creates a new CP/M system disk.
.bp
.ll 65
.in 0
.ce
.sh
Table 1-2.  (continued)
.qs
.sp
.ll 60
.in 5
.nf
Command                      Function
.fi
.sp
.in 16
.ti -11
SUBMIT     Submits a file of commands for batch processing.
.sp
.ti -11
DUMP       Dumps the contents of a file in hex.
.sp
.ti -11
MOVCPM     Regenerates the CP/M system for a particular memory size.
.sp
.ll 65
.in 0
.pp
Transient commands are specified in the same manner as built-in commands, and
additional commands are easily defined by the user.  For convenience, the
transient command can be preceded by a drive name which causes the transient
to be loaded from the specified drive into the TPA for execution.  Thus, the
command
.sp
.ti 8
B:STAT
.sp
causes CP/M to temporarily log in drive B for the source of the STAT
transient, and then return to the original logged disk for subsequent
processing.
.sp 2
.tc         1.6.1  STAT Command
.sh
1.6.1  STAT Command
.qs
.sp
.ul
Syntax:
.qu
.sp
.in 8
.nf
STAT
STAT "command line"
.fi
.in 0
.pp
The STAT command provides general statistical information about file storage
and device assignment.  Special forms of the command line allow the current
device assignment to be
examined and altered.  The various command lines that can be specified are
shown with an explanation of each form to the right.
.sp 2
.in 24
.ti -16
STAT            If you type an empty command line, the STAT transient
calculates the storage remaining on all active drives, and prints 
one of the following messages:
.sp
.nf
d: R/W, SPACE:  nnnK
.sp
d: R/O, SPACE:  nnnK
.fi
.sp
for each active drive d:, where R/W indicates the drive can be read or
written, and R/O indicates the drive is Read-Only (a drive becomes R/O by
explicitly setting it to Read-Only, as shown below, or by inadvertently
changing disks without performing a warm start).  The space remaining on
the disk in drive d: is given in kilobytes by nnn.
.sp
.ti -16
STAT d:         If a drive name is given, then the drive is selected before
the storage is computed.  Thus, the command STAT B: could be issued while
logged into drive A, resulting in the message
.sp
BYTES REMAINING ON B:  nnnK
.sp
.ti -16
STAT afn        The command line can also specify a set of files to be
scanned by STAT.  The files that satisfy afn are listed in alphabetical
order, with storage requirements for each file under the heading:
.sp
.nf
RECS BYTES EXT D:FILENAME.TYP
rrrr bbbK ee d:filename.typ
.fi
.sp
where rrrr is the number of 128-byte records allocated to the file, bbb is
the number of kilobytes allocated to the file (bbb=rrrr*128/1024), ee is the
number of 16K extensions (ee=bbb/16), d is the drive name containing the
file (A...P), filename is the eight-character primary filename, and
typ is the three-character filetype.  After listing the individual
files, the storage usage is summarized.
.sp
.ti -16
STAT d:afn      The drive name can be given ahead of the afn.  The specified
drive is first selected, and the form STAT afn is executed.
.sp
.ti -16
STAT d:=R/O     This form sets the drive given by d to Read-Only, remaining
in effect until the next warm or cold start takes place.  When a disk is
Read-Only, the message
.sp
BDOS ERR ON d:  Read-Only
.sp
appears if there is an attempt to write to the Read-Only disk.  CP/M
waits until a key is pressed before performing an automatic 
warm start, at
which time the disk becomes R/W.
.in 0
.bp
.pp
The STAT command allows you to control the physical-to-logical device
assignment.  See the IOBYTE function described in Sections 5 and 6.  There
are four logical peripheral devices that are, at any particular instant, each
assigned one of several physical peripheral devices.  The 
following is a list of the four logical devices:
.sp 2
.in 5
.ti -2
o CON: is the system console device, used by CCP for communication with the
operator.
.sp
.ti -2
o RDR: is the paper tape reader device.
.sp
.ti -2
o PUN: is the paper tape punch device.
.sp
.ti -2
o LST: is the output list device.
.in 0
.sp
.pp
The actual devices attached to any particular computer system are driven by
subroutines in the BIOS portion of CP/M.  Thus, the logical RDR: device, for
example, could actually be a high speed reader, teletype reader, or cassette
tape.  To allow some flexibility in device naming and assignment, several
physical devices are defined in Table 1-3.
.sp 2
.ce
.sh
Table 1-3.  Physical Devices
.ll 60
.in 5
.sp
.nf
Device                       Meaning
.fi
.sp
.in 14
.ti -8
TTY:    Teletype device (slow speed console)
.sp
.ti -8
CRT:    Cathode ray tube device (high speed console)
.sp
.ti -8
BAT:    Batch processing (console is current RDR:, output goes to current
LST: device)
.sp
.ti -8
UC1:    User-defined console
.sp
.ti -8
PTR:    Paper tape reader (high speed reader)
.sp
.ti -8
UR1:    User-defined reader #1
.sp
.ti -8
UR2:    User-defined reader #2
.sp
.ti -8
PTP:    Paper tape punch (high speed punch)
.sp
.ti -8
UP1:    User-defined punch #1
.sp
.ti -8
UP2:    User-defined punch #2
.sp
.ti -8
LPT:    Line printer
.sp
.ti -8
UL1:    User-defined list device #1
.in 0
.ll 65
.nx oneb


