.bp 1
.op
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.he
.ft                                3-%
.pc 1
.tc 3  CP/M Assembler
.ce 2
.sh
Section 3
.sp
.sh
CP/M Assembler
.qs
.sp 3
.tc    3.1  Introduction
.he CP/M Operating System Manual                    3.1  Introduction
.sh
3.1  Introduction
.qs
.pp 5
The CP/M assembler reads assembly-language source files from the 
disk and produces 8080 machine language in Intel hex format.  
To start the CP/M assembler, type a command in one of the 
following forms:
.sp
.nf
.in 8
ASM filename
ASM filename.parms
.fi
.in 0
.sp
In both cases, the assembler assumes there is a file on the
disk with the name:
.sp
.ti 8
filename.ASM
.sp
which contains an 8080 assembly-language source file.  The first
and second forms shown above differ only in that the second form
allows parameters to be passed to the assembler to control source
file access and hex and print file destinations.
.pp
In either case, the CP/M assembler loads and prints the message:
.sp
.ti 8
CP/M ASSEMBLER VER n.n
.sp
where n.n is the current version number.  In the case of the 
first command, the assembler reads the source file with assumed 
filetype ASM and creates two output files
.sp
.in 8
.nf
filename.HEX
filename.PRN
.fi
.in 0
.pp
The HEX file contains the machine code corresponding to the 
original program in Intel hex format, and the PRN file contains 
an annotated listing showing generated machine code, error flags, 
and source lines.  If errors occur during translation, they are
listed in the PRN file and at the console.
.pp
The form ASM filename parms is used to redirect input and 
output files from their defaults.  In this case, the parms 
portion of the command is a three-letter group that specifies the 
origin of the source file, the destination of the hex file, and 
the destination of the print file.  The form is
.bp
.ti 8
filename.p1p2p3
.sp
where p1, p2, and p3 are single letters.  P1 can be
.sp
.ti 8
A,B, ...,P
.sp
which designates the disk name that contains the source file.  P2 
can be
.sp
.ti 8
A,B, ...,P
.sp
which designates the disk name that will receive the hex file; or, P2 
can be 
.sp
.ti 8
Z
.sp
which skips the generation of the hex file.
.pp
P3 can be
.sp
.ti 8
A,B, ...,P
.sp
which designates the disk name that will receive the 
print file.  P3 can also be specified as
.sp
.ti 8
X
.sp
which places the listing at the console; or
.sp
.ti 8
Z
.sp
which skips generation of the print file.  Thus, the command
.sp
.ti 8
ASM X.AAA
.sp
indicates that the source, X.HEX, and print, X.PRN, files 
are also to be created on disk A.  This form of the command is 
implied if the assembler is run from disk A.  Given that you are 
currently addressing disk A, the above command is the same as
.sp
.ti 8
ASM X
.sp
The command
.sp
.ti 8
ASM X.ABX
.sp
indicates that the source file is to be taken from disk A, the 
hex file is to be placed on disk B, and the listing file is to be 
sent to the console.  The command
.sp
.ti 8
ASM X.BZZ
.sp
takes the source file from disk B and skips the generation of the 
hex and print files.  This command is useful for fast execution of 
the assembler to check program syntax.
.bp
.pp
The source program format is compatible with the Intel 8080 
assembler.  Macros are not implemented in ASM; see the optional 
MAC macro assembler.  There are certain extensions in the CP/M 
assembler that make it somewhat easier to use.  These extensions 
are described below.
.sp 2
.tc    3.2  Program Format
.he CP/M Operating System Manual                  3.2  Program Format
.sh
3.2  Program Format
.qs
.pp
An assembly-language program acceptable as input to the assembler 
consists of a sequence of statements of the form
.sp
.ti 8
line# label operation operand ;comment
.sp
where any or all of the fields may be present in a particular 
instance.  Each assembly-language statement is terminated with a 
carriage return and line-feed (the line-feed is inserted 
automatically by the ED program), or with the character !, which 
is treated as an end-of-line by the assembler.  Thus, multiple 
assembly-language statements can be written on the same physical 
line if separated by exclamation point symbols.
.pp
The line# is an optional decimal integer value representing the 
source program line number, and ASM ignores this field if 
present.
.pp
The label field takes either of the following forms:
.sp
.in 8
.nf
identifier
identifier:
.fi
.in 0
.sp
The label field is optional, except where noted in particular statement 
types.  The identifier is a sequence of alphanumeric characters 
where the first character is alphabetic.  Identifiers can be 
freely used by the programmer to label elements such as program 
steps and assembler directives, but cannot exceed 16 characters 
in length.  All characters are significant in an identifier, 
except for the embedded dollar symbol $, which can be used to 
improve readability of the name.  Further, all lower-case 
alphabetics are treated as upper-case.  The 
following are all valid instances of labels:
.sp 2
.nf
.in 8
x       xy      long$name

x:      yxl:    longer$named$data:

X1Y2    X1x2    x234$5678$9012$3456:
.fi
.in 0
.sp
.pp
The operation field contains either an assembler directive or 
pseudo operation, or an 8080 machine operation code.  The pseudo 
operations and machine operation codes are described in Section 
3.3.
.pp
Generally, the operand field of the statement contains an 
expression formed out of constants and labels, along with 
arithmetic and logical operations on these elements.  Again, the 
complete details of properly formed expressions are given in 
Section 3.3.
.pp
The comment field contains arbitrary characters following the 
semicolon  
symbol until the next real or logical end-of-line.  These 
characters are read, listed, and otherwise ignored by the 
assembler.  The CP/M assembler also treats statements that begin 
with an * in column one as comment statements that are listed 
and ignored in the assembly process.
.pp
The assembly-language program is formulated as a sequence of 
statements of the above form, terminated by an optional END 
statement.  All statements following the END are ignored by the 
assembler.
.sp 2
.tc    3.3  Forming the Operand
.he CP/M Operating System Manual             3.3  Forming the Operand
.sh
3.3  Forming the Operand
.qs
.pp
To describe the operation codes and pseudo operations completely, 
it is necessary first to present the form of the operand field, 
since it is used in nearly all statements.  Expressions in the 
operand field consist of simple operands, labels, constants, and 
reserved words, combined in properly formed subexpressions by 
arithmetic and logical operators.  The expression computation is 
carried out by the assembler as the assembly proceeds.  Each 
expression must produce a 16-bit value during the assembly.  
Further, the number of significant digits in the result must not 
exceed the intended use.  If an expression is to be used 
in a byte move immediate instruction, the most significant 8 bits 
of the expression must be zero.  The restriction on the 
expression significance is given with the individual 
instructions.
.sp 2
.tc         3.3.1  Labels
.sh
3.3.1  Labels
.qs
.pp
As discussed above, a label is an identifier that occurs on a 
particular statement.  In general, the label is given a value 
determined by the type of statement that it precedes.  If the 
label occurs on a statement that generates machine code or 
reserves memory space (for example, a MOV instruction or a
DS pseudo operation), the label is given the value of the program address 
that it labels.  If the label precedes an EQU or SET, the label 
is given the value that results from evaluating the operand 
field.  Except for the SET statement, an identifier can label 
only one statement.
.pp
When a label appears in the operand field, its value is 
substituted by the assembler.  This value can then be combined 
with other operands and operators to form the operand field for a 
particular instruction.
.bp
.tc         3.3.2  Numeric Constants
.sh
3.3.2  Numeric Constants
.qs
.pp
A numeric constant is a 16-bit value in one of several bases.  
The base, called the radix of the constant, is denoted by a 
trailing radix indicator.  The following are radix indicators:
.sp
.in 3
.nf
o B is a binary constant (base 2).
o O is a octal constant (base 8).
o Q is a octal constant (base 8).
o D is a decimal constant (base 10).
o H is a hexadecimal constant (base 16).
.fi
.in 0
.pp
Q is an alternate radix indicator for octal numbers because the 
letter O is easily confused with the digit 0.  Any numeric 
constant that does not terminate with a radix indicator is 
a decimal constant.
.pp
A constant is composed as a sequence of digits, followed by 
an optional radix indicator, where the digits are in the 
appropriate range for the radix.  Binary constants must 
be composed of 0 and 1 digits, octal constants can contain digits 
in the range 0-7, while decimal constants contain decimal digits.  
Hexadecimal constants contain decimal digits as well as 
hexadecimal digits A(10D), B(11D), C(12D), D(13D), E(14D), and 
F(15D).  Note that the leading digit of a 
hexadecimal constant must be a decimal digit to avoid confusing a 
hexadecimal constant with an identifier.  A leading 0 will always 
suffice.  A constant composed in this manner must evaluate to a 
binary number that can be contained within a 16-bit counter, 
otherwise it is truncated on the right by the assembler.
.pp
Similar 
to identifiers, embedded $ signs are allowed within constants to 
improve their readability.  Finally, the radix indicator is 
translated to upper-case if a lower-case letter is encountered.  
The following are all valid instances of numeric constants:
.sp 2
.nf
.in 8
1234      1234D     1100B     1111$0000$1111$0000B
.sp
1234H     OFFEH     3377O     33$77$22Q
.sp
3377o     Ofe3h     1234d     Offffh
.fi
.in 0
.sp 2
.tc         3.3.3  Reserved Words
.sh
3.3.3  Reserved Words
.qs
.pp
There are several reserved character sequences that have 
predefined meanings in the operand field of a statement.  The 
names of 8080 registers are given below.  When they are 
encountered, they produce the values shown to the right.
.bp
.nf
.sh
                 Table 3-1.  Reserved Characters
.sp
                       Character     Value

                           A           7
                           B           0
                           C           1
                           D           2
                           E           3
                           H           4
                           L           5
                           M           6
                           SP          6
                           PSW         6
.fi
.in 0
.sp
.pp
Again, lower-case names have the same values as their upper-case 
equivalents.  Machine instructions can also be used in the 
operand field; they evaluate to their internal codes. In the case 
of instructions that require operands, where the specific operand 
becomes a part of the binary bit pattern of the instruction, 
for example, MOV A,B, the value of the instruction, in this case MOV, 
is the bit pattern of the instruction with zeros in the optional 
fields, for example, MOV produces 40H.
.pp
When the symbol $ occurs in the operand field, not embedded 
within identifiers and numeric constants, its value becomes the 
address of the next instruction to generate, not including the 
instruction contained within the current logical line.
.sp 2
.tc         3.3.4  String Constants
.sh
3.3.4  String Constants
.qs
.pp
String constants represent sequences of ASCII characters and are 
represented by enclosing the characters within apostrophe symbols.
All strings must be fully contained within the current 
physical line (thus allowing exclamation point symbols within 
strings) and must 
not exceed 64 characters in length.  The apostrophe character 
itself can be included within a string by representing it as a 
double apostrophe (the two keystrokes ''), which becomes a single 
apostrophe when read by the assembler.  In most cases, the string 
length is restricted to either one or two characters (the DB 
pseudo operation is an exception), in which case the string 
becomes an 8- or 16-bit value, respectively.  Two-character 
strings become a 16-bit constant, with the second character as 
the low-order byte, and the first character as the high-order 
byte.
.pp
The value of a character is its corresponding ASCII code.  There 
is no case translation within strings; both upper- and 
lower-case characters can be represented.  You should note 
that only graphic printing ASCII characters are 
allowed within strings.
.bp
.nf
           Valid strings:          How assembler reads strings:

     'A' 'AB' 'ab' 'c'               A  AB  ab  c
     '' 'a''' '''' ''''                 a'  '   '
     'Walla Walla Wash.'             Walla Walla Wash.
     'She said ''Hello'' to me.'     She said ''Hello'' to me
     'I said ''Hello'' to her.'      I said ''Hello'' to her
.fi
.sp 2
.tc         3.3.5  Arithmetic and Logical Operators
.sh
3.3.5  Arithmetic and Logical Operators
.qs
.pp
The operands described in Section 3.3 can be combined in normal algebraic 
notation using any combination of properly formed operands, 
operators, and parenthesized expressions.  The operators 
recognized in the operand field are described in Table 3-2.
.sp 2
.ce
.sh
Table 3-2.  Arithmetic and Logical Operators
.ll 60
.in 5
.sp
.nf
Operators                         Meaning
.sp
.fi
.in 19
.ti -13
a + b        unsigned arithmetic sum of a and b
.sp
.ti -13
a - b        unsigned arithmetic difference between a and b
.sp
.ti -13
  + b        unary plus (produces b)
.sp
.ti -13
  - b        unary minus (identical to 0 - b)
.sp
.ti -13
a * b        unsigned magnitude multiplication of a and b
.sp
.ti -13
a / b        unsigned magnitude division of a by b
.sp
.ti -13
a MOD b      remainder after a / b.
.sp
.ti -13
NOT b        logical inverse of b (all 0s become 1s, 1s become 
0s), where b is considered a 16-bit value
.sp
.ti -13
a AND b      bit-by-bit logical and of a and b
.sp
.ti -13
a OR b       bit-by-bit logical or of a and b
.sp
.ti -13
a XOR b      bit-by-bit logical exclusive or of a and b
.sp
.ti -13
a SHL b      the value that results from shifting a to the left 
by an amount b, with zero fill
.sp
.ti -13
a SHR b      the value that results from shifting a to the 
right by an amount b, with zero fill
.in 0
.ll 65
.sp
.pp
In each case, a and b represent simple operands (labels, numeric 
constants, reserved words, and one- or two-character strings) or 
fully enclosed parenthesized subexpressions, like those shown in 
the following examples:
.sp 2
.nf
.in 8
10+20  10h+37Q  LI/3  (L2+4) SHR 3

('a' and 5fh) + '0' ('B'+B) OR (PSW+M)

(1+(2+c)) shr (A-(B+1))
.fi
.in 0
.sp
.pp
Note that all computations are performed at assembly time as 16-bit
unsigned operations.  Thus, -1 is computed as 0-1, which 
results in the value 0ffffh (that is, all 1s).  The resulting 
expression must fit the operation code in which it is used.  For 
example, if the expression is used in an ADI (add immediate)
instruction, the high-order 8 bits of the expression must be 
zero.  As a result, the operation ADI-1 produces an error message 
(-1 becomes 0ffffh, which cannot be represented as an 8-bit 
value), while ADI(-1) AND 0FFH is accepted by the assembler 
because the AND operation zeros the high-order bits of the 
expression.
.sp 2
.tc         3.3.6  Precedence of Operators
.sh
3.3.6  Precedence of Operators
.qs
.pp
As a convenience to the programmer, ASM assumes that operators 
have a relative precedence of application that allows the 
programmer to write expressions without nested levels of 
parentheses.  The resulting expression has assumed parentheses 
that are defined by the relative precedence.  The order of 
application of operators in unparenthesized expressions is listed 
below.  Operators listed first have highest precedence (they are 
applied first in an unparenthesized expression), while operators 
listed last have lowest precedence.  Operators listed on the same 
line have equal precedence, and are applied from left to right as 
they are encountered in an expression.
.sp 2
.in 8
.mb 5
.fm 1
.nf
* / MOD SHL SHR

- +

NOT

AND

OR XOR
.fi
.in 0
.sp
.pp
Thus, the expressions shown to the left below are interpreted by 
the assembler as the fully parenthesized expressions shown to the 
right.
.bp
.nf
.in 5
a*b+c                       (a*b)+c

a+b*c                       a+(b*c)

a MOD b*c SHL d             ((a MOD b)*c) SHL d

a OR b AND NOT c+d SHL e    a OR (b AND (NOT (c+(d SHL e))))
.fi
.in 0
.sp
.pp
Balanced, parenthesized subexpressions can always be used to 
override the assumed parentheses; thus, the last expression above 
could be rewritten to force application of operators in a 
different order, as shown:
.sp
.ti 8
(a OR b) AND (NOT c)+ d SHL e
.sp
This results in these assumed parentheses:
.sp
.ti 8
(a OR b) AND ((NOT c) + (d SHL e))
.pp
An unparenthesized expression is well-formed only if the 
expression that results from inserting the assumed parentheses is 
well-formed.
.sp 2
.tc    3.4  Assembler Directives
.he CP/M Operating System Guide             3.4  Assembler Directives
.sh
3.4  Assembler Directives
.qs
.pp
Assembler directives are used to set labels to specific values 
during the assembly, perform conditional assembly, define storage 
areas, and specify starting addresses in the program.  Each 
assembler directive is denoted by a pseudo operation that appears 
in the operation field of the line.  The acceptable pseudo operations
are shown in Table 3-3.
.sp 2
.nf
.sh
                Table 3-3.  Assembler Directives
.sp
        Directive                 Meaning
.fi
.sp
          ORG       set the program or data origin
.sp
          END       end program, optional start address
.sp
          EQU       numeric equate
.sp
          SET       numeric set
.sp
          IF        begin conditional assembly
.sp
          ENDIF     end of conditional assembly
.sp
          DB        define data bytes
.sp
          DW        define data words
.sp
          DS        define data storage area
.in 0
.bp
.tc         3.4.1  The ORG Directive
.sh
3.4.1  The ORG Directive
.qs
.pp
The ORG statement takes the form:
.sp
.ti 8
label  ORG  expression
.sp
where label is an optional program identifier and expression is 
a 16-bit expression, consisting of operands that are defined 
before the ORG statement.  The assembler begins machine code 
generation at the location specified in the expression.  There 
can be any number of ORG statements within a particular program, 
and there are no checks to ensure that the programmer is not 
defining overlapping memory areas.  Note that 
most programs written for the CP/M system begin with an ORG 
statement of the form:
.sp
.ti 8
ORG  100H
.sp
which causes machine code generation to begin at the base of the 
CP/M transient program area.  If a label is specified in the ORG 
statement, the label is given the value of the expression.  This 
label can then be used in the operand field of other statements 
to represent this expression.
.sp 2
.tc         3.4.2  The END Directive
.sh
3.4.2  The END Directive
.qs
.pp
The END statement is optional in an assembly-language program, 
but if it is present it must be the last statement.  All 
subsequent statements are ignored in the assembly.  The END 
statement takes the following two forms:
.sp
.in 8
.nf
label END

label END expression
.fi
.in 0
.sp
where the label is again optional.  If the first form is used, 
the assembly process stops, and the default starting address of 
the program is taken as 0000.  Otherwise, the expression is 
evaluated, and becomes the program starting address.  This 
starting address is included in the last record of the Intel-formatted
machine code hex file that results from the 
assembly.  Thus, most CP/M assembly-language programs end with 
the statement:
.sp
.ti 8
END  100H
.sp
resulting in the default starting address of 100H (beginning of 
the transient program area).
.bp
.tc         3.4.3  The EQU Directive
.sh
3.4.3  The EQU Directive
.qs
.pp
The EQU (equate) statement is used to set up synonyms for 
particular numeric values.  The EQU statement takes the form:
.sp
.ti 8
.nf
label    EQU    expression
.fi
.sp
where the label must be present and must not label any other 
statement.  The assembler evaluates the expression and assigns 
this value to the identifier given in the label field.  The 
identifier is usually a name that describes the value in a more 
human-oriented manner.  Further, this name is used throughout the 
program to place parameters on certain functions.  Suppose data 
received from a teletype appears on a particular input port, and 
data is sent to the teletype through the next output port in 
sequence.  For example, you can use this series of equate statements
to define these ports for a particular hardware environment:
.sp 2
.in 8
.nf
TTYBASE      EQU 10H         ;BASE PORT NUMBER FOR TTY

TTYIN        EQU TTYBASE     ;TTY DATA IN

TTYOUT       EQU TTYBASE+1   ;TTY DATA OUT
.fi
.in 0
.sp
.pp
At a later point in the program, the statements that access the 
teletype can appear as follows:
.sp 2
.in 7
.nf
 IN     TTYIN     ;READ TTY DATA TO REG-A

 ...

 OUT    TTYOUT    ;WRITE DATA TO TTY FROM REG-A
.fi
.in 0
.sp 2
making the program more readable than if the absolute I/O ports 
are used.  Further, if the hardware environment is redefined 
to start the teletype communications ports at 7FH instead of 10H, 
the first statement need only be changed to
.sp
.ti 8
.nf
TTYBASE    EQU    7FH    ;BASE PORT NUMBER FOR TTY
.fi
.sp
and the program can be reassembled without changing any other 
statements.
.sp 2
.tc         3.4.4  The SET Directive
.sh
3.4.4  The SET Directive
.qs
.pp
The SET statement is similar to the EQU, taking the form:
.sp
.ti 8
label    SET    expression
.sp
except that the label can occur on other SET statements within 
the program.  The expression is evaluated and becomes the current 
value associated with the label.  Thus, the EQU statement defines 
a label with a single value, while the SET statement defines a 
value that is valid from the current SET statement to the point 
where the label occurs on the next SET statement.  The use of the 
SET is similar to the EQU statement, but is used most often in 
controlling conditional assembly.
.sp 2
.tc         3.4.5  The IF and ENDIF Directives
.sh
3.4.5  The IF and ENDIF Directives
.qs
.pp
The IF and ENDIF statements define a range of assembly-language 
statements that are to be included or excluded during the 
assembly process.  These statements take on the form:
.sp 2
.in 8
.nf
IF  expression

statement#1

statement#2

  ...

statement#n

ENDIF
.fi
.in 0
.sp
.pp
When encountering the IF statement, the assembler evaluates the 
expression following the IF.  All operands in the expression must 
be defined ahead of the IF statement.  If the expression 
evaluates to a nonzero value, then statement#1 through 
statement#n are assembled.  If the expression evaluates to zero, 
the statements are listed but not assembled.  Conditional 
assembly is often used to write a single generic program that 
includes a number of possible run-time environments, with only a 
few specific portions of the program selected for any particular 
assembly.  The following program segments, for example, might be 
part of a program that communicates with either a teletype or a 
CRT console (but not both) by selecting a particular value for 
TTY before the assembly begins.
.bp
.nf
.in 8
TRUE     EQU            OFFFFH     ;DEFINE VALUE OF TRUE
FALSE    EQU            NOT TRUE   ;DEFINE VALUE OF FALSE
;
TTY      EQU            TRUE       ;TRUE IF TTY, FALSE IF CRT
;
TTYBASE  EQU            10H        ;BASE OF TTY I/O PORTS
CRTBASE  EQU            20H        ;BASE OF CRT I/O PORTS
         IF             TTY        ;ASSEMBLE RELATIVE TO
                                   ;TTYBASE
CONIN    EQU            TTYBASE    ;CONSOLE INPUT
CONOUT   EQU            TTYBASE+1  ;CONSOLE OUTPUT
         ENDIF

;        IF             NOT TTY    ;ASSEMBLE RELATIVE TO
                                   ;CRTBASE
CONIN    EQU            CRTBASE    ;CONSOLE INPUT
CONOUT   EQU            CRTBASE+1  ;CONSOLE OUTPUT

         ENDIF
         ...
         IN             CONIN      ;READ CONSOLE DATA
         ...
         OUT            CONTOUT    ;WRITE CONSOLE DATA
.fi
.in 0
.sp 2
In this case, the program assembles for an environment where 
a teletype is connected, based at port 10H.  The statement 
defining TTY can be changed to
.sp
.nf
.ti 8
TTY      EQU         FALSE
.fi
.sp
and, in this case, the program assembles for a CRT based at 
port 20H.
.sp 2
.tc         3.4.6  The DB Directive
.sh
3.4.6  The DB Directive
.qs
.pp
The DB directive allows the programmer to define initialized 
storage areas in single-precision byte format.  The DB statement 
takes the form:
.sp
.nf
.ti 8
label DB e#1, e#2, ..., e#n
.fi
.sp
where e#1 through e#n are either expressions that evaluate to 8-bit
values (the high-order bit must be zero) or are ASCII strings 
of length no greater than 64 characters.  There is no practical 
restriction on the number of expressions included on a single 
source line.  The expressions are evaluated and placed 
sequentially into the machine code file following the last 
program address generated by the assembler.  String characters 
are similarly placed into memory starting with the first 
character and ending with the last character.  Strings of length 
greater than two characters cannot be used as operands in more 
complicated expressions.
.bp
.sh
Note:  \c
.qs
ASCII 
characters are always placed in memory with the parity bit reset 
(0).  Also, there is no translation from lower- to upper-case 
within strings.  The optional label can be used to reference the 
data area throughout the remainder of the program.  The following 
are examples of valid DB statements:
.sp 2
.nf
.in 8
data:        DB        0,1,2,3,4,5
             DB        data and 0ffh,5,377Q,1+2+3+4

sign-on:     DB        'please type your name',cr,lf,0
             DB        'AB' SHR 8, 'C', 'DE' AND 7FH
.fi
.in 0
.sp 3
.tc         3.4.7  The DW Directive
.sh
3.4.7  The DW Directive
.qs
.pp
The DW statement is similar to the DB statement except double-precision
two-byte words of storage are initialized.  The DW statement 
takes the form:
.sp
.nf
.ti 8
label        DW        e#1, e#2, ..., e#n
.fi
.sp
where e#1 through e#n are expressions that evaluate to 16-bit 
results.  Note that ASCII strings of one or two 
characters are allowed, but strings longer than two characters 
are disallowed.  In all cases, the data storage is consistent 
with the 8080 processor; the least significant byte of the 
expression is stored first in memory, followed by the most 
significant byte.  The following are examples of DW statements:
.sp 2
.in 8
.nf
doub:        DW        0ffefh,doub+4,signon-$,255+255
             DW        'a', 5, 'ab', 'CD', 6 shl 8 or llb.
.fi
.in 0
.sp 3
.tc         3.4.8  The DS Directive
.sh
3.4.8  The DS Directive
.qs
.pp
The DS statement is used to reserve an area of uninitialized 
memory, and takes the form:
.sp
.ti 8
.nf
label        DS        expression
.fi
.sp
where the label is optional.  The assembler begins subsequent 
code generation after the area reserved by the DS.  Thus, the DS 
statement given above has exactly the same effect as the following 
statement:
.sp
.nf
.in 7
label:       EQU  $  ;LABEL VALUE IS CURRENT CODE LOCATION
             ORG  $+expression  ;MOVE PAST RESERVED AREA
.fi
.in 0
.nx threeb





